### 2016-02-01 mandatory_assignment/main.tex
started working on assignment

\documentclass{amsart}

\usepackage[]{enumerate} 
\usepackage[]{pst-platon} 
\usepackage[]{graphicx}

\def\Gobble#1{}

\theoremstyle{plain}
\newtheorem{theorem}{Theorem}
\theoremstyle{definition}
\newtheorem{definition}{Definition}

\title{Euler's polyhedron formula}
\author{Ivar Haugal{\o}kken Stangeby}
\date{\today}

\begin{document}

\begin{abstract}
   In this short paper we look at Euler's polyhedron formula. This formula was
   one of my first experiences with more abstract mathematics, and I have loved
   it ever since. The proof is very simple, and extremely elegant.
\end{abstract}
\maketitle

\tableofcontents

\section{Introduction and motivation}
\label{sec:introduction_and_motivation}

As a vivid reader of popular science one often comes across the problem
regarding the \emph{Seven Bridges of K{\"o}ningsberg}. This city, located in
what was then known as Prussia had a river running through the middle. This
river naturally divided the city into the mainland --- on either side --- as
well as two large islands. Connecting these islands with the mainland were
seven bridges. This is where the famous question arises:

\begin{quote}
    \emph{Can the seven bridges of the city of K{\"o}ningsberg, over the river
    Preger, all be traversed in a single trip without doubling back and have
the trip end in the same place it began?}  
\end{quote}

Eminent mathematician Leonhard Euler proved that the problem stated above has
no solution. In working with the solution to this problem however, he dabbled
in a type of mathematics later to be known as \emph{graph theory}. One of the
more remarkable results (in my humble opinion) to arise from this new field of
mathematics is the discovery of a relationship between the number of vertices,
edges and faces in a polyhedron. It is exactly this relationship that Euler's
polyhedron formula captures.

\begin{theorem}[Euler's polyhedron formula]
    Any convex polyhedron's surface satisfies the equation
    \begin{equation}
        \notag
        V - E + F = 2
    \end{equation}
    where $V$, $E$ and $F$ denote the number of vertices, edges and faces,
    respectively.
\end{theorem}
\section{Preliminaries}
\label{sec:preliminaries}

In order to get a firm grasp of this formula, we first need to recall some
elementary geometry.  We start with the definition of the \emph{polyhedron}.

\begin{definition}[Polyhedron]
    A \emph{polyhedron} is a solid in three dimensions with flat polygonal
    faces, straight edges and sharp corners or vertices.\footnote{This
    definition is not very precice, but it will suffice for our purposes.}
\end{definition}

Polyhedra are named based on the number of faces using the greek prefix system.
Some examples include the \emph{hexahedron}, otherwise known as the cube with
its six faces; the \emph{octahedron} or \emph{double pyramid} with its eight
faces; and the \emph{icosahedron} with its 20 faces.

An important tool when studying platonic solids is that of the \emph{Schlegel
diagram}. It is essentially a way of projecting a $d$-dimensional object into
$d-1$ dimensions. In our case, we want to study the three dimensional platonic
solids in the two dimensional plane. Each platonic solid, when projected into
two dimensions, form what we call a \emph{planar graph}.

\begin{definition}[Planar graph]
  A \emph{planar graph}, is a graph that can be embedded in the plane, i.e., it
  can be drawn on the plane in such a way that its edges intersect only at
  their endpoints.
\end{definition}

\section{Proof of the formula}
\label{sec:proof_of_the_formula}

The proof of Euler's polyhedron formula is very simple and applies to any
convex polyhedron. We will use a cube as our example. We first take note of the
number of vertices, faces and edges. A cube has six faces, $F = 6$, twelve
edges, $E = 12$, and eight vertices, $V = 8$. This yields
\begin{equation}
    \label{eq:form}
    V - E + F = 8 - 12 + 6 = 2.
\end{equation}

Our goal now is to progressively remove edges, faces and vertices until we are
left with a very simple polygon. We will use transformations that preserve the
relation shown in equation \ref{eq:form}. We start by projecting the polyhedron
into the plane using a Schlegel diagram. This is done by removing one face and
squishing the cube into the plane, as shown in the leftmost image in figure
\ref{fig:proof}. Note that removing one face does not alter the number of
vertices and edges, hence we now have $V - E + F = 1$.

We now triangulate any face with more than three edges until all faces are
triangular. This is done by adding one edge, but doing so also adds a new face.
Hence the quantity $V - E + F$ is unchanged.

\begin{figure}[ht!]
    \centering
    \includegraphics{drawing.pdf}
    \caption{First iterations of proof; using a cube.}
    \label{fig:proof}
\end{figure}

We now have two possible transformations to iterate: 
\begin{enumerate}
    \item Remove a triangle with one edge adjacent to the exterior as shown in
        the middle image in figure \ref{fig:proof}.  
    \item Remove a triangle
        with two edges adjacent to the exterior as shown in the rightmost image
        in figure \ref{fig:proof}.
\end{enumerate}
Note that neither of these operations change the quantity of interest. One
removes one edge and one face and the other removes one vertex, two edges, and
one face. We apply these transformations while always keeping the requirement
that the outer edges form a simple cycle. We will eventually end up with a
triangle with three vertices, three edges and one face.

\end{document}

### 2016-02-10 mandatory_assignment/main.tex
meeting notes

@@
    one of my first experiences with more abstract mathematics, and I have loved
-   it ever since. The proof is very simple, and extremely elegant.
+   it ever since. The proof we consider here is very simple, however there is a
+   surprising amount of different proofs for this same result. \cite{proofs}
+   presents a list of 20 of them.
 \end{abstract}
@@
 that the outer edges form a simple cycle. We will eventually end up with a
-triangle with three vertices, three edges and one face.
+triangle with three vertices, three edges and one face. This final shape
+clearly satisfies ${V - E + F = 1}$.  Adding the face we removed when producing
+the Schlegel-diagram gives us ${V - E + F = 2}$.
 
+\begin{thebibliography}{9}
+    \bibitem{proofs}
+        http://www.ics.uci.edu/~eppstein/junkyard/euler/all.html
+\end{thebibliography}
 \end{document}

### 2016-02-10 project/visualization/src/barth_sextic.py
first iteration plotting

@@ -1,0 +1,11 @@
+from implicit_plot import implicit_plot
+import mayavi.mlab as mlab
+
+figw = mlab.figure(1, bgcolor=(0.1, 0.1, 0.1), size=(400, 400))
+phi = 1.618
+phi_squared = 1.618**2
+
+implicit_plot('4*(%f*x**2 - y**2)*(%f*y**2-z**2)*(%f*z**2-x**2) - (1 + 2*%f)*(x**2 + y**2 + z**2 - 1)**2' % (phi_squared, phi_squared, phi_squared, phi),
+              (-2, 2, -2, 2, -2, 2), col_osurf=(1.0,0.086,1.0),
+              fig_handle=figw, opa_val=1.0, opaque=True, ori_axis=True)
+mlab.show()

### 2016-02-10 project/visualization/src/cube.py

@@ -1,0 +1,7 @@
+from implicit_plot import implicit_plot
+import mayavi.mlab as mlab
+figw = mlab.figure(1, bgcolor=(0.1, 0.1, 0.1), size=(400, 400))
+implicit_plot('x**20 + y**20 + z**20 - 100',
+              (-2, 2, -2, 2, -2, 2), col_osurf=(0.87,0.086,0.086),
+              fig_handle=figw, opa_val=1.0, opaque=True, ori_axis=False)
+mlab.show()

### 2016-02-10 project/visualization/src/implicit_plot.py

@@ -1,0 +1,97 @@
+import numpy as np
+import mayavi.mlab as mlab
+def implicit_plot(expr, ext_grid, fig_handle=None, Nx=101, Ny=101, Nz=101,
+                 col_isurf=(50/255, 199/255, 152/255), col_osurf=(240/255,36/255,87/255),
+                 opa_val=0.8, opaque=True, ori_axis=True, **kwargs):
+    """Function to plot algebraic surfaces described by implicit equations in Mayavi
+ 
+    Implicit functions are functions of the form
+ 
+        `F(x,y,z) = c`
+ 
+    where `c` is an arbitrary constant.
+ 
+    Parameters
+    ----------
+    expr : string
+        The expression `F(x,y,z) - c`; e.g. to plot a unit sphere, the `expr` will be `x**2 + y**2 + z**2 - 1`
+    ext_grid : 6-tuple
+        Tuple denoting the range of `x`, `y` and `z` for grid; it has the form - (xmin, xmax, ymin, ymax, zmin, zmax)
+    fig_handle : figure handle (optional)
+        If a mayavi figure object is passed, then the surface shall be added to the scene in the given figure. Then, it is the responsibility of the calling function to call mlab.show().
+    Nx, Ny, Nz : Integers (optional, preferably odd integers)
+        Number of points along each axis. It is recommended to use odd numbers to ensure the calculation of the function at the origin.
+    col_isurf : 3-tuple (optional)
+        color of inner surface, when double-layered surface is used. This is also the specified color for single-layered surface.
+    col_osurf : 3-tuple (optional)
+        color of outer surface
+    opa_val : float (optional)
+        Opacity value (alpha) to use for surface
+    opaque : boolean (optional)
+        Flag to specify whether the surface should be opaque or not
+    ori_axis : boolean
+        Flag to specify whether a central axis to draw or not
+ 
+    """
+    if fig_handle==None:  # create a new figure
+        fig = mlab.figure(1,bgcolor=(0.97, 0.97, 0.97), fgcolor=(0, 0, 0), size=(800, 800))
+    else:
+        fig = fig_handle
+    xl, xr, yl, yr, zl, zr = ext_grid
+    x, y, z = np.mgrid[xl:xr:eval('{}j'.format(Nx)),
+                       yl:yr:eval('{}j'.format(Ny)),
+                       zl:zr:eval('{}j'.format(Nz))]
+    scalars = eval(expr)
[…]

### 2016-02-10 project/visualization/src/sphere.py

@@ -1,0 +1,8 @@
+import mayavi.mlab as mlab
+import numpy as np
+
+from implicit_plot import implicit_plot
+
+figw = mlab.figure(1, bgcolor=(0.1, 0.1, 0.1), size=(400, 400))
+implicit_plot('x**2 + y**2 + z**2 - {R:f}**2'.format(R=1), (-3, 3, -3, 3, -3, 3), fig_handle=figw, Nx=64, Ny=64, Nz=64, col_isurf=(0.0,0.2,0.8), opaque=True, ori_axis=False)
+mlab.show()

### 2016-02-10 project/visualization/src/barth_sextic.py
fixed interior color

@@ -6,6 +6,6 @@ phi = 1.618
 phi_squared = 1.618**2
 
 implicit_plot('4*(%f*x**2 - y**2)*(%f*y**2-z**2)*(%f*z**2-x**2) - (1 + 2*%f)*(x**2 + y**2 + z**2 - 1)**2' % (phi_squared, phi_squared, phi_squared, phi),
-              (-2, 2, -2, 2, -2, 2), col_osurf=(1.0,0.086,1.0),
+              (-2, 2, -2, 2, -2, 2), col_osurf=(1.0,0.086,1.0), col_isurf=(0.7, 0.1, 0.7),
               fig_handle=figw, opa_val=1.0, opaque=True, ori_axis=True)
 mlab.show()

### 2016-02-10 project/visualization/src/implicit_plot.py

@@ -1,5 +1,6 @@
 import numpy as np
 import mayavi.mlab as mlab
+
 def implicit_plot(expr, ext_grid, fig_handle=None, Nx=101, Ny=101, Nz=101,
                  col_isurf=(50/255, 199/255, 152/255), col_osurf=(240/255,36/255,87/255),
                  opa_val=0.8, opaque=True, ori_axis=True, **kwargs):

### 2016-02-10 project/visualization/src/barth_sextic.py
updated notes

@@ -9,3 +9,4 @@ implicit_plot('4*(%f*x**2 - y**2)*(%f*y**2-z**2)*(%f*z**2-x**2) - (1 + 2*%f)*(x*
               (-2, 2, -2, 2, -2, 2), col_osurf=(1.0,0.086,1.0), col_isurf=(0.7, 0.1, 0.7),
               fig_handle=figw, opa_val=1.0, opaque=True, ori_axis=True)
 mlab.show()
+mlab.savefig('test.pdf')

### 2016-02-11 project/visualization/src/klein.py
added klein bottle

@@ -1,0 +1,7 @@
+from implicit_plot import implicit_plot
+import mayavi.mlab as mlab
+figw = mlab.figure(1, bgcolor=(0.1, 0.1, 0.1), size=(400, 400))
+implicit_plot('(x**2 + y**2 + z**2 + 2*y - 1) * ((x**2 + y**2 + z**2 -2*y - 1)**2 - 8*z**2) + 16*x*z*(x**2 + y**2 + z**2 - 2*y - 1)',
+              (-5, 5, -5, 5, -5, 5), col_osurf=(0.87,0.086,0.086), col_isurf=(1.00, 1.00, 1.00),
+              fig_handle=figw, opa_val=1.0, opaque=True, ori_axis=False)
+mlab.show()

### 2016-02-11 project/visualization/src/birational_transformation.py
added a general birational_transformation function

@@ -1,0 +1,22 @@
+def quadratic_transform(expr, iterations=1, x='(y*z)', y='(x*z)', z='(x*y)'):
+    """
+    Maps (x, y, z) to (yz, xz, xy)
+    Optional argument, number of iterations of map
+    """
+    if iterations == 0:
+        return expr
+
+    for i in range(iterations):
+        new_string = ""
+        for char in expr:
+            if char == 'x':
+                new_string += x
+            elif char == 'y':
+                new_string += y
+            elif char == 'z':
+                new_string += z
+            else:
+                new_string += char
+            expr = new_string
+        
+    return new_string

### 2016-02-19 project/article/src/article.tex
Started working on projective spaces

\documentclass{article}

% mathematical typesetting
\usepackage[]{amsmath} 
\usepackage[]{amsthm} 
\usepackage[]{amssymb}

% lists
\usepackage[]{enumerate}

% bibliography management
\usepackage[backend=biber, style=alphabetic]{biblatex} 
\addbibresource{../bib/mat2000.bib}

% index management
\usepackage[]{makeidx} 
\makeindex

% cross-references
\usepackage[]{hyperref} 
\usepackage[]{cleveref}

% proclamation definitions
\theoremstyle{definition}
\newtheorem{defn}{Definition}
\newtheorem{exm}{Example}
\theoremstyle{plain}
\newtheorem{thm}{Theorem}
\newtheorem{lma}{Lemma}
\newtheorem{crl}{Corollary}
\newtheorem{prp}{Proposition}

% auxilliary commands
\newcommand{\proj}{\ensuremath{\mathbb{P}}} % projective space
\newcommand{\R}{\ensuremath{\mathbb{R}}}    % the real numbers
\newcommand{\projp}[3]{\ensuremath{[ #1 : #2 : #3 ]}}

\title{Visualizing Algebraic Surfaces}
\author{Ivar Haugal{\o}kken Stangeby}

\begin{document}

\maketitle

\begin{abstract}

\end{abstract}

\section{Introduction}
\label{sec:introduction}

\section{Projective Space}
\label{sec:projective_space}

We start our study of algebraic surfaces by looking at \emph{projective space}.
Quite informally, projective space formalizes the notion of parallell lines
intersecting at infinity. In order to get a better understanding of the general
projective space $\proj^n$ we will first construct the \emph{real projective
plane}\index{real projective plane} $\proj^2(\R)$. According to \cite{Wik16}
there are three equivalent definitions:

\begin{enumerate}[I.]
    \item\label{def:one} The set of all lines in $\R^3$ passing through the origin $\left( 0,
        0, 0 \right)$. Every such line meets the \emph{sphere} of radius one
        centered in the origin exactly twice, say in $P = (x, y, z)$ and its
        antipodal point $(-x, -y, -z)$.
    \item\label{def:two} The points on the sphere $S^2$, where every point $P$ and its
        antipodal points are not distinguished. For example, the point $(1, 0,
        0)$ is identified with the point $(-1, 0, 0)$.
    \item The set of equivalence classes of $\R^3 \setminus \left\{(0, 0, 0)\right\}$ where
        two points $P$ and $P^\prime$ are considered equivalent if and only if
        there is a positive $\lambda \in \R$ such that $P = \lambda P^\prime$.
\end{enumerate}

The last definition is the one seemingly most commonly used, and is therefore
the one we will employ here. Note that the elements in the real projective
plane are equivalence classes of points in $\R^3$. We denote an element in
$\proj^2(\R)$ as $\projp{x}{y}{z}$. These elements are commonly referred to as
\emph{homogenous coordinates}\index{homogenous coordinates}. We state this
formally in a definition.

\begin{defn}[The real projective plane]
    Let $\sim$ be the equivalence relation on $\R^3$ defined by
    \begin{equation}
        \notag
        (x, y, z) \sim (x', y', z') \iff (x, y, z) = \lambda(x, y, z),
    \end{equation}
    where $\lambda$ is some positive real number. Then we denote the
    equivalence class of $(x, y, z)$ as $\projp{x}{y}{z}$ and we define
    \emph{the real projective plane} as the set
    \begin{equation}
        \notag
        \proj^2(\R) = \left\{ \projp{x}{y}{z} \mid (x, y, z) \in \R^3 \setminus
        \left\{ (0, 0, 0) \right\} \right\}.
    \end{equation}
\end{defn}

We can also look at a subset of $\proj^2(\R)$, namely the set of points
$\projp{x}{y}{z}$ where $z \neq 0$. Remembering the identification from the
above definition, these are equivalent to $\projp{\frac{x}{z}}{\frac{y}{z}}{1}$
with $\lambda = 1 / z$. If we consider definition \ref{def:two} from above,
this corresponds to the set of points on the northern hemisphere of $S^2$, not
including the circle of intersection in the $(x,y)$-plane. Similarly, we can
look at the set of points $\projp{x}{y}{0}$. This corresponds, in $S^2$, to the
circle of intersection in the $(x, y)$-plane.

\section{Algebraic Surfaces}
\label{sec:algebraic_surfaces}

\printindex
\printbibliography

\end{document}

### 2016-02-19 project/article/src/article.tex
added a front page image and def of projective space

@@
 
+% image handling
+\usepackage[]{graphicx}
+
 % proclamation definitions
@@
 
+\begin{titlepage}
+
 \maketitle
+
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=0.8\linewidth]{../pictures/labs_septic.png}
+\end{figure}
+
+\end{titlepage}
 
@@
 \end{abstract}
+
+\tableofcontents
 
@@
 projective space $\proj^n$ we will first construct the \emph{real projective
-plane}\index{real projective plane} $\proj^2(\R)$. According to \cite{Wik16}
-there are three equivalent definitions:
+plane}\index{real projective plane} $\proj^2(\R)$.
+
+\subsection{The real projective plane}
+According to \cite{Wik16} there are three equivalent definitions:
 
@@
 
+We can now take the notion of the projective plane and generalize to $n$
+dimensions.
+\subsection{Projective Space}
+\label{sub:projective_space}
+
+The general definition of projective $n$-space is completely anologous to the
+definition of the projective plane, the set of lines in $\R^{n+1}$ passing
+through the origin, however we include a formal definition for completeness.
+
+\begin{defn}[The real projective space of dimension $n$]\index{projective space}
+    Let $\sim$ be an equivalence relation on $\R^{n+1}\setminus \left\{ \left(
+    0, \ldots, 0 \right) \right\}$ defined by
+    \begin{equation}
+        \notag
+        \left( x_1, \ldots, x_{n+1} \right) \sim \left( x_1', \ldots, x_{n+1}' \right) 
+        \iff \left( x_1, \ldots, x_{n+1} \right) = \lambda \left( x_1', \ldots, x_{n+1}' \right)
+    \end{equation} 
+    where $\lambda$ is some positive real number. We then denote the
+    equivalence class of $\left( x_1, \ldots, x_{n+1} \right)$ by
+    $\projp{x_1}{\ldots}{x_{n+1}}$ and define \emph{the real projective space
+    of dimension $n$} as the set
+    \begin{equation}
+        \notag
+        \proj^n(\R) = \left\{ \projp{x_1}{\ldots}{x_{n+1}} \mid \left( x_1, \ldots, x_{n+1} \right) \in \R^{n+1} \setminus \left\{ \left( 0, \ldots, 0 \right) \right\} \right\}.
+    \end{equation}
+
+\end{defn}
+
 \section{Algebraic Surfaces}

### 2016-02-25 mandatory_assignment/main.tex
assignment done

@@
 \usepackage[]{graphicx} 
+\usepackage[]{url} 
+\urlstyle{sf}
 
@@
 in a type of mathematics later to be known as \emph{graph theory}. One of the
-more remarkable results (in my humble opinion) to arise from this new field of
+more remarkable results (in my opinion) to arise from this new field of
 mathematics is the discovery of a relationship between the number of vertices,
@@
 \end{theorem}
+
+The reason I chose this specific theorem is because it was one of my first
+encounters with something resembling pure mathematics than the kind of
+mathematics I had been exposed to during elementary school. Having always been
+a fan of puzzles, I especially liked this result since it sprung from a
+seemingly ''innocent'' looking question that everyday people would walk around
+pondering.
+
 \section{Preliminaries}
@@
     \includegraphics{drawing.pdf}
-    \caption{First iterations of proof; using a cube.}
+    \caption{First iterations of proof; using a cube. Image retrieved from \cite{image}.}
     \label{fig:proof}
@@
 
-\begin{thebibliography}{9}
+\section{Applications}
+\label{sec:applications}
+
+Direct applications of the polyhedron formula can be hard to find. However, its
+generalizations to other mathematical fields as topology gives us an indication
+as to what kind of topological object we are working with. The \emph{euler
+characteristic}, denoted $\chi$, is a topological invariant. In this more
+general sense, our polyhedron formula reads $\chi = V - E + F$, where $\chi =
+2$. This characteristic can for instance be used to classifly the platonic
+solids. \cite{chi} presents a lot of mathematical results using the euler
+characteristic.
+
+\begin{thebibliography}{2}
     \bibitem{proofs}
-        http://www.ics.uci.edu/~eppstein/junkyard/euler/all.html
+        \url{http://www.ics.uci.edu/~eppstein/junkyard/euler/all.html}
+    \bibitem{image}
+        \url{https://en.wikipedia.org/wiki/Euler_characteristic#/media/File:V-E%2BF%3D2_Proof_Illustration.svg}
+    \bibitem{chi}
+        \url{https://en.wikipedia.org/wiki/Euler_characteristic}
 \end{thebibliography}

### 2016-02-25 mandatory_assignment/main.tex
spelling mistakes

@@
     faces, straight edges and sharp corners or vertices.\footnote{This
-    definition is not very precice, but it will suffice for our purposes.}
+    definition is not very precise, but it will suffice for our purposes.}
 \end{definition}
 
-Polyhedra are named based on the number of faces using the greek prefix system.
+Polyhedra are named based on the number of faces using the Greek prefix system.
 Some examples include the \emph{hexahedron}, otherwise known as the cube with
@@
 general sense, our polyhedron formula reads $\chi = V - E + F$, where $\chi =
-2$. This characteristic can for instance be used to classifly the platonic
-solids. \cite{chi} presents a lot of mathematical results using the euler
+2$. This characteristic can for instance be used to classify the platonic
+solids. \cite{chi} presents a lot of mathematical results using the Euler
 characteristic.

### 2016-04-01 project/article/src/article.tex
backup

@@
 \usepackage[backend=biber, style=alphabetic]{biblatex} 
+\usepackage[]{url} 
+\urlstyle{sf}
 \addbibresource{../bib/mat2000.bib}
@@
         0)$ is identified with the point $(-1, 0, 0)$.
-    \item The set of equivalence classes of $\R^3 \setminus \left\{(0, 0, 0)\right\}$ where
+    \item The set of equivalence classes of $\R^3 \setminus (0, 0, 0)$ where
         two points $P$ and $P^\prime$ are considered equivalent if and only if
@@
 
-\begin{defn}[The real projective plane]
+\begin{defn}[The real projective plane]\index{$\proj^2(\R)$}
     Let $\sim$ be the equivalence relation on $\R^3$ defined by
@@
         \proj^2(\R) = \left\{ \projp{x}{y}{z} \mid (x, y, z) \in \R^3 \setminus
-        \left\{ (0, 0, 0) \right\} \right\}.
+        (0, 0, 0) \right\}.
     \end{equation}
@@
 
-\begin{defn}[The real projective space of dimension $n$]\index{projective space}
-    Let $\sim$ be an equivalence relation on $\R^{n+1}\setminus \left\{ \left(
-    0, \ldots, 0 \right) \right\}$ defined by
+\begin{defn}[The real projective space of dimension $n$]\index{projective
+    space}\index{$\proj^n(\R)$}
+
+    Let $\sim$ be an equivalence relation on $\R^{n+1}\setminus \left( 0,
+    \ldots, 0 \right)$ defined by
     \begin{equation}
@@
         \notag
-        \proj^n(\R) = \left\{ \projp{x_1}{\ldots}{x_{n+1}} \mid \left( x_1, \ldots, x_{n+1} \right) \in \R^{n+1} \setminus \left\{ \left( 0, \ldots, 0 \right) \right\} \right\}.
+        \proj^n(\R) = \left\{ \projp{x_1}{\ldots}{x_{n+1}} \mid \left( x_1, \ldots, x_{n+1} \right) 
+        \in \R^{n+1} \setminus \left( 0, \ldots, 0 \right) \right\}.
     \end{equation}
-
 \end{defn}
@@
 
+Algebraic surfaces will be the central object of study in this project. We
+follow the definition of \cite{cite}.
+
 \printindex

### 2016-04-01 project/visualization/src/birational_transform.py

@@ -1,8 +1,9 @@
-def quadratic_transform(expr, iterations=1, x='(y*z)', y='(x*z)', z='(x*y)'):
+def birational_transform(expr, iterations=1, x='(y*z)', y='(x*z)', z='(x*y)'):
     """
     Maps (x, y, z) to (yz, xz, xy)
     Optional argument, number of iterations of map
     """
+
     if iterations == 0:
         return expr
 

### 2016-04-01 project/visualization/src/implicit_plotter.py

@@ -1,0 +1,23 @@
+import numpy as np
+import mayavi.mlab as mlab
+
+def implicit_plotter(expr, ext_grid, grid_res=(101, 101, 101),**kwargs):
+    """
+    Parameters
+    ----------
+    expr: string
+        A string representation of 'f(x, y, z) - c = 0'
+    ext_grid: 6-tuple
+        The discrete grid '(xmin, xmax, ymin, ymax, zmin, zmax)'
+    """
+        
+    # Create a new figure
+    fig = mlab.figure(1, bgcolor=(1.0, 1.0, 1.0), fgcolor=(0, 0, 0), size=(800, 800))
+    xl, xr, yl, yr, zl, zr = ext_grid
+    Nx, Ny, Nz = grid_res
+    x, y, z = np.mgrid[xl:xr:eval('{}j'.format(Nx)),
+                       yl:yr:eval('{}j'.format(Ny)),
+                       zl:zr:eval('{}j'.format(Nz))]
+
+if __name__ == "__main__":
+    implicit_plotter(expr='x + y + z', ext_grid=(0, 1, 0, 1, 0, 1))

### 2016-04-14 project/article/src/article.tex
finished informal introduction, draft

--- 2016-02-19 project/article/src/article.tex
@@
-\documentclass{article}
+\documentclass[a4paper]{article}
 
-% mathematical typesetting
 \usepackage[]{amsmath} 
@@
 
-% lists
-\usepackage[]{enumerate}
+\usepackage[demo]{graphicx} 
+\usepackage[]{caption} 
+\usepackage[]{subcaption}
 
-% bibliography management
-\usepackage[backend=biber, style=alphabetic]{biblatex} 
-\addbibresource{../bib/mat2000.bib}
-
-% index management
-\usepackage[]{makeidx} 
-\makeindex
-
-% cross-references
-\usepackage[]{hyperref} 
 \usepackage[]{cleveref}
 
-% proclamation definitions
-\theoremstyle{definition}
-\newtheorem{defn}{Definition}
-\newtheorem{exm}{Example}
-\theoremstyle{plain}
-\newtheorem{thm}{Theorem}
-\newtheorem{lma}{Lemma}
-\newtheorem{crl}{Corollary}
-\newtheorem{prp}{Proposition}
-
-% auxilliary commands
-\newcommand{\proj}{\ensuremath{\mathbb{P}}} % projective space
-\newcommand{\R}{\ensuremath{\mathbb{R}}}    % the real numbers
-\newcommand{\projp}[3]{\ensuremath{[ #1 : #2 : #3 ]}}
+\newcommand{\C}{\ensuremath{\mathbb{C}}}
 
@@
 \author{Ivar Haugal{\o}kken Stangeby}
+\date{\today}
 
@@
     \maketitle
+    \tableofcontents
 
-\begin{abstract}
+    \section{An Informal Introduction}
+    \label{sec:an_informal_introduction}
 
-\end{abstract}
+    In this section we briefly look at the natural construction of the
+    mathematical objects we are interested in studying in the rest of this
+    paper.
 
-\section{Introduction}
-\label{sec:introduction}
+    \subsection{Real Algebraic Curves}
+    \label{sub:real_algebraic_curves}
 
-\section{Projective Space}
-\label{sec:projective_space}
+    From elementary mathematics one learns about real valued functions, $f(x)$,
+    and how to graph these functions by setting $y = f(x)$ and plotting points
+    in the $(x, y)$-plane. Now, the \emph{graph of a function} is something of
+    a peculiarity, because it comes with some restrictions. Not all curves in
+    the $(x, y)$-plane correspond to functions. The method for verifying
+    whether a certain ''graph'' corresponds to a function or not typically
+    taught in school is the \emph{vertical line test}.
 
-We start our study of algebraic surfaces by looking at \emph{projective space}.
-Quite informally, projective space formalizes the notion of parallell lines
-intersecting at infinity. In order to get a better understanding of the general
-projective space $\proj^n$ we will first construct the \emph{real projective
-plane}\index{real projective plane} $\proj^2(\R)$. According to \cite{Wik16}
-there are three equivalent definitions:
+    Having an equation $y = f(x)$ we can form what we call the \emph{equation
+    of a curve at zero}. We define a new function in two variables
+    \begin{equation}
+        \label{eq:equation_at_zero}
+        g(x, y) = y - f(x) = 0.
+    \end{equation}
+    If the function $f(x)$ is a polynomial in the variable $x$ with certain
+    coefficients we call the function graph of $f(x)$ \emph{algebraic}.
 
-\begin{enumerate}[I.]
-    \item\label{def:one} The set of all lines in $\R^3$ passing through the origin $\left( 0,
-        0, 0 \right)$. Every such line meets the \emph{sphere} of radius one
-        centered in the origin exactly twice, say in $P = (x, y, z)$ and its
-        antipodal point $(-x, -y, -z)$.
-    \item\label{def:two} The points on the sphere $S^2$, where every point $P$ and its
-        antipodal points are not distinguished. For example, the point $(1, 0,
-        0)$ is identified with the point $(-1, 0, 0)$.
-    \item The set of equivalence classes of $\R^3 \setminus \left\{(0, 0, 0)\right\}$ where
-        two points $P$ and $P^\prime$ are considered equivalent if and only if
-        there is a positive $\lambda \in \R$ such that $P = \lambda P^\prime$.
-\end{enumerate}
+    Generally speaking, we call a curve defined by \cref{eq:equation_at_zero}
+    an \emph{algebraic curve} if the function $g(x, y)$ is a polynomial in two
+    variables, $x$ and $y$. Mathematically, this can be expressed as
+    \begin{equation}
+        \label{eq:algebraic_curve}
+        g(x, y) = \sum^{}_{i, j} a_{i_j}x^{i}y^{i}.
+    \end{equation}
 
-The last definition is the one seemingly most commonly used, and is therefore
-the one we will employ here. Note that the elements in the real projective
-plane are equivalence classes of points in $\R^3$. We denote an element in
-$\proj^2(\R)$ as $\projp{x}{y}{z}$. These elements are commonly referred to as
-\emph{homogenous coordinates}\index{homogenous coordinates}. We state this
-formally in a definition.
+    However, if this distinction between a graph and a curve is to be justified
+    there has to be some curves that are not graphs. One of the first examples
+    one encounter of a curve not corresponding to a function is what you get
+    when you set $y^2 = x^3$. This curve does not pass the vertical-line-test
+    and is therefore something different from a graph. Similarly, the equation
+    $y^2 = x^3 - x^2$ defines a curve that again is not a graph. These are
+    shown in \cref{fig:curves_not_graphs}. These curves exhibit \emph{singular
+    points} at $(0, 0)$.
 
-\begin{defn}[The real projective plane]
-    Let $\sim$ be the equivalence relation on $\R^3$ defined by
+    \begin{figure}[h]
+        \centering
+        \begin{subfigure}{0.5\textwidth}
+            \centering
+            \includegraphics[width=0.35\linewidth]{foo}
+            \caption{$g(x, y) = y^2 - x^3 = 0$}
+        \end{subfigure}%
+        \begin{subfigure}{0.5\textwidth}
+            \centering
+            \includegraphics[width=0.35\linewidth]{bar}
+            \caption{$g(x, y) = y^2 - x^3 + x^2 = 0$}
+        \end{subfigure} 
+        \caption{Examples of curves that are \emph{not} graphs given by their
+        functions $g(x, y)$}.}
+        \label{fig:curves_not_graphs}
+    \end{figure}
+
+    With what we have so far, we can move up a dimension. Instead of
+    considering curves in the $(x, y)$-plane we can look at surfaces in the
+    $(x, y, z)$-space. We, as we did with $g(x, y)$ above, define a function
+    $h(x, y, z) = z - g(x, y)$. The surfaces are given by equations on the form
+    $h(x, y, z) = 0$. These surfaces are called algebraic if $h(x, y, z)$ is a
+    polynomial in the variables $x, y, z$. Again, mathematically, this is
+    expressed as
     \begin{equation}
-        \notag
-        (x, y, z) \sim (x', y', z') \iff (x, y, z) = \lambda(x, y, z),
+        \label{eq:algebraic_surfaces}
+        h(x, y, z) = \sum^{}_{i, j, k} a_{ijk}x^iy^jz^k.
     \end{equation}
-    where $\lambda$ is some positive real number. Then we denote the
-    equivalence class of $(x, y, z)$ as $\projp{x}{y}{z}$ and we define
-    \emph{the real projective plane} as the set
+
+    \subsection{Going Complex}
+    \label{sub:going_complex}
+
+    The surfaces considered so far are over the real numbers, i.e, the
+    coefficients $a_{ijk}$ are real numbers. We can instead work over the
+    complex numbers where the surfaces are given by an equivalent equation, but
+    where the coefficients now are complex numbers\footnote{We will come back
+    to why we make this transition when we talk about sets being algebraically
+    closed or not.}:
     \begin{equation}
-        \notag
-        \proj^2(\R) = \left\{ \projp{x}{y}{z} \mid (x, y, z) \in \R^3 \setminus
-        \left\{ (0, 0, 0) \right\} \right\}.
+        \left\{ (x, y, z) \in \C \mid h(x, y, z) = 0 \right\}.
     \end{equation}
-\end{defn}
+    Again, doing the same trick (one-trick ponies) we can now define $w = h(x,
+    y, z)$ and look at the equations $w - h(x, y, z) = 0$. Now, we are stuck
+    with an equation in four variables. With only three degrees of freedom to
+    work with when visualizing these objects we encounter an obstacle. How do
+    we know what these surfaces look like? This brings us to the world of
+    \emph{projective geometry}.
 
-We can also look at a subset of $\proj^2(\R)$, namely the set of points
-$\projp{x}{y}{z}$ where $z \neq 0$. Remembering the identification from the
-above definition, these are equivalent to $\projp{\frac{x}{z}}{\frac{y}{z}}{1}$
-with $\lambda = 1 / z$. If we consider definition \ref{def:two} from above,
-this corresponds to the set of points on the northern hemisphere of $S^2$, not
-including the circle of intersection in the $(x,y)$-plane. Similarly, we can
-look at the set of points $\projp{x}{y}{0}$. This corresponds, in $S^2$, to the
-circle of intersection in the $(x, y)$-plane.
-
-\section{Algebraic Surfaces}
-\label{sec:algebraic_surfaces}
-
-\printindex
-\printbibliography
+    \section{Projective Geometry}
+    \label{sec:projective_geometry}
 

### 2016-04-14 project/article/src/article.tex
moved over projective geom stuff

@@
 
+    Quite informally, projective space formalizes the notion of parallell lines
+    intersecting at infinity. In order to get a better understanding of the
+    general projective space $\proj^n$ we will first construct the \emph{real
+    projective plane}\index{real projective plane} $\proj^2(\R)$.
+
+\subsection{The real projective plane}
+
+    According to \cite{Wik16} there are three equivalent definitions:
+    \begin{enumerate}[I.]
+        \item\label{def:one} The set of all lines in $\R^3$ passing through the origin $\left( 0,
+            0, 0 \right)$. Every such line meets the \emph{sphere} of radius one
+            centered in the origin exactly twice, say in $P = (x, y, z)$ and its
+            antipodal point $(-x, -y, -z)$.
+        \item\label{def:two} The points on the sphere $S^2$, where every point $P$ and its
+            antipodal points are not distinguished. For example, the point $(1, 0,
+            0)$ is identified with the point $(-1, 0, 0)$.
+        \item The set of equivalence classes of $\R^3 \setminus (0, 0, 0)$ where
+            two points $P$ and $P^\prime$ are considered equivalent if and only if
+            there is a positive $\lambda \in \R$ such that $P = \lambda P^\prime$.
+    \end{enumerate}
+
+    The last definition is the one seemingly most commonly used, and is therefore
+    the one we will employ here. Note that the elements in the real projective
+    plane are equivalence classes of points in $\R^3$. We denote an element in
+    $\proj^2(\R)$ as $\projp{x}{y}{z}$. These elements are commonly referred to as
+    \emph{homogenous coordinates}\index{homogenous coordinates}. We state this
+    formally in a definition.
+
+    \begin{defn}[The real projective plane]\index{$\proj^2(\R)$}
+        Let $\sim$ be the equivalence relation on $\R^3$ defined by
+        \begin{equation}
+            \notag
+            (x, y, z) \sim (x', y', z') \iff (x, y, z) = \lambda(x, y, z),
+        \end{equation}
+        where $\lambda$ is some positive real number. Then we denote the
+        equivalence class of $(x, y, z)$ as $\projp{x}{y}{z}$ and we define
+        \emph{the real projective plane} as the set
+        \begin{equation}
+            \notag
+            \proj^2(\R) = \left\{ \projp{x}{y}{z} \mid (x, y, z) \in \R^3 \setminus
+            (0, 0, 0) \right\}.
+        \end{equation}
+    \end{defn}
+
+    We can also look at a subset of $\proj^2(\R)$, namely the set of points
+    $\projp{x}{y}{z}$ where $z \neq 0$. Remembering the identification from the
+    above definition, these are equivalent to $\projp{\frac{x}{z}}{\frac{y}{z}}{1}$
+    with $\lambda = 1 / z$. If we consider definition \ref{def:two} from above,
+    this corresponds to the set of points on the northern hemisphere of $S^2$, not
+    including the circle of intersection in the $(x,y)$-plane. Similarly, we can
+    look at the set of points $\projp{x}{y}{0}$. This corresponds, in $S^2$, to the
+    circle of intersection in the $(x, y)$-plane.
+
+    We can now take the notion of the projective plane and generalize to $n$
+    dimensions.
+    \subsection{Projective Space}
+    \label{sub:projective_space}
+
+    The general definition of projective $n$-space is completely anologous to the
+    definition of the projective plane, the set of lines in $\R^{n+1}$ passing
+    through the origin, however we include a formal definition for completeness.
+
+    \begin{defn}[The real projective space of dimension $n$]\index{projective
+        space}\index{$\proj^n(\R)$}
+
+        Let $\sim$ be an equivalence relation on $\R^{n+1}\setminus \left( 0,
+        \ldots, 0 \right)$ defined by
+        \begin{equation}
+            \notag
+            \left( x_1, \ldots, x_{n+1} \right) \sim \left( x_1', \ldots, x_{n+1}' \right) 
+            \iff \left( x_1, \ldots, x_{n+1} \right) = \lambda \left( x_1', \ldots, x_{n+1}' \right)
+        \end{equation} 
+        where $\lambda$ is some positive real number. We then denote the
+        equivalence class of $\left( x_1, \ldots, x_{n+1} \right)$ by
+        $\projp{x_1}{\ldots}{x_{n+1}}$ and define \emph{the real projective space
+        of dimension $n$} as the set
+        \begin{equation}
+            \notag
+            \proj^n(\R) = \left\{ \projp{x_1}{\ldots}{x_{n+1}} \mid \left( x_1, \ldots, x_{n+1} \right) 
+            \in \R^{n+1} \setminus \left( 0, \ldots, 0 \right) \right\}.
+        \end{equation}
+    \end{defn}
+
 \end{document}

### 2016-04-14 project/article/src/article.tex
pulled over the stuff on projective space, generalized to complex setting

--- 2016-04-14 project/article/src/article.tex
@@
 
+\usepackage[backend=biber, style=alphabetic]{biblatex} 
+\addbibresource{../bib/mat2000.bib}
+
 \usepackage[demo]{graphicx} 
@@
 
+\usepackage[]{enumerate}
+
 \usepackage[]{cleveref}
 
+\theoremstyle{definition}
+\newtheorem{defn}{Definition}
+
 \newcommand{\C}{\ensuremath{\mathbb{C}}}
+\newcommand{\proj}{\ensuremath{\mathbb{P}}}
 
@@
 
+    The way we visualize these surfaces in four variables is by \emph{fixing}
+    one of the variables. This is called \emph{homogenizing} the polynomial
+    $h(x, y, z)$. We first want to introduce some definitions.
+
+    Quite informally, projective space formalizes the notion of parallell lines
+    intersecting at infinity. In order to get a better understanding of the
+    general projective space $\proj^n$ we will first construct the
+    \emph{projective plane} $\proj^2(\C)$.
+
+    \subsection{The Projective Plane}
+    \label{sub:the_projective_plane}
+
+    According to \cite{Wik16} there are three equivalent definitions:
+    \begin{enumerate}
+        \item The set of all lines in $\C^3$ passing through the origin $(0, 0,
+            0)$. Every such line meets the \emph{sphere} of radius one centered
+            in the origin exactly twice, say in $P = (x, y, z)$ and its
+            antipodal point $(-x, -y, -z)$.
+
+        \item\label{item:sphere} The points on the sphere $S^2$, where every point $P$ and its
+            antipodal points are not distinguished. For example, the point $(1,
+            0, 0)$ is identified with the point $(-1, 0, 0)$.
+
+        \item The set of equivalence classes of $\C^3 \setminus \left\{ 0
+            \right\}$ where two points $P$ and $P'$ are considered equivalent
+            if and only if there is a non-zero $\lambda \in \C$ such that $P =
+            \lambda P'$.
+    \end{enumerate}
+
+    The last definition is the one seemingly most used, and is therefore the
+    one we will employ here. Note that the elements in the projective plane are
+    equivalence classes of points in $\C^3$. We denote an element in
+    $\proj^2(\C)$ as $\left[ x : y : z \right]$. These elements are commonly
+    referred to as \emph{homogenous coordinates}. We state this formally in a
+    definition.
+
+    \begin{defn}[The projective plane]
+        Let $\sim$ be the equivalence relation on $\C^3$ defined by
+        \begin{equation}
+            \notag
+            (x, y, z) \sim (x', y', z') \iff (x', y', z') = \lambda(x, y, z),
+        \end{equation}
+        where $\lambda$ is some non-zero scalar in $\C$. We denote the
+        equivalence class of $(x, y, z)$ as $\left[ x : y : z \right]$ and we
+        define the \emph{projective plane} as the set
+        \begin{equation}
+            \notag
+            \proj^2(\C) = \left\{ \left[ x : y : z \right] \mid (x, y, z) \in
+            \C^3 \setminus \left\{ 0 \right\}\right\}.
+        \end{equation}
+    \end{defn}
+
+    We can now look at an interesting subset of $\proj^2(\C)$, namely the set
+    of points $\left[x : y : z \right]$ where $z \neq 0$. Remembering the
+    identification from the above definition, these are equivalent to $\left[
+    \frac{x}{z} : \frac{y}{z} : 1 \right]$ where $\lambda = 1 / z$. If we
+    consider definition \cref{item:sphere} from above, this corresponds to the
+    set of points on the northern hemisphere of $S^2$ not including the circle
+    of intersection in the $(x, y)$-plane. Similarly, we can look at the set of
+    points $\left[ x : y : 0 \right]$. This correponds, in $S^2$, to the circle
+    of intersection in the $(x, y)$-plane.
+
+    We can now take the notion of the
+    projective plane and generalize this to $n$ dimensions.
+
+    \subsection{Projective Space}
+    \label{sub:projective_space}
+
+    The general definition of projective $n$-space is completely anologous to
+    the definition of the projective plane, the set of lines in $\C^{n+1}$
+    passing through the origin, however we include a formal definition for
+    completeness.
+
+    \begin{defn}[The projective space of dimension $n$]
+        Let $\sim$ be an equivalence relation on $\C^{n+1} \setminus \left\{ 0
+        \right\}$ defined by
+        \begin{equation}
+            \notag
+            \left( x_1, \ldots, x_{n+1} \right) \sim \left( x_1', \ldots,
+            \x_{n+1}' \right) \iff \left( x_1', \ldots, x_{n+1}' \right) =
+            \lambda \left( x_1, \ldots, x_{n+1} \right)
+        \end{equation}
+        where $\lambda$ is some non-zero complex number. We then denote the
+        equivalence class of $\left( x_1, \ldots, x_{n+1} \right)$ by $\left[
+        x_1 : \ldots : x_{n+1} \right]$ and define the \emph{projective space
+        of dimension $n$} as the set
+        \begin{equation}
+            \notag
+            \proj^n{\C} = \left\{ \left[ x_1 : \ldots : x_{n+1} \right] \mid
+            \left( x_1, \ldots, x_{n+1} \right) \in \C^{n+1} \setminus \left\{
+            0 \right\} \right\}.
+        \end{equation}
+    \end{defn}
+
+\printbibliography
+
 \end{document}

### 2016-04-22 project/article/src/article.tex
typing out a s-storm

@@
 \addbibresource{../bib/mat2000.bib}
+\usepackage[]{url} 
+\urlstyle{sf}
 
-\usepackage[demo]{graphicx} 
+\usepackage[]{graphicx} 
+\usepackage[]{wrapfig} 
 \usepackage[]{caption} 
@@
 \newtheorem{defn}{Definition}
+\newtheorem{exmpl}{Example}
+
+\theoremstyle{plain}
+\newtheorem{lma}{Lemma}
+\newtheorem{crl}{Corollary}
 
@@
 \newcommand{\proj}{\ensuremath{\mathbb{P}}}
+\renewcommand{\hom}{\ensuremath{h_{\mathrm{hom}}}}
 
@@
 
+    \begin{titlepage}
     \maketitle
+
+    \begin{figure}[htb]
+        \centering
+        \includegraphics[width=0.8\linewidth]{../pictures/labs_septic.png}
+    \end{figure}
+    \end{titlepage}
     \tableofcontents
@@
             \centering
-            \includegraphics[width=0.35\linewidth]{foo}
+            \includegraphics[width=1.0\linewidth]{../pictures/cusp.pdf}
             \caption{$g(x, y) = y^2 - x^3 = 0$}
@@
             \centering
-            \includegraphics[width=0.35\linewidth]{bar}
+            \includegraphics[width=1.0\linewidth]{../pictures/double_point.pdf}
             \caption{$g(x, y) = y^2 - x^3 + x^2 = 0$}
@@
     Again, doing the same trick (one-trick ponies) we can now define $w = h(x,
-    y, z)$ and look at the equations $w - h(x, y, z) = 0$. Now, we are stuck
-    with an equation in four variables. With only three degrees of freedom to
-    work with when visualizing these objects we encounter an obstacle. How do
-    we know what these surfaces look like? This brings us to the world of
-    \emph{projective geometry}.
+    y, z)$ and look at the equations $w - h(x, y, z) = 0$ in the $(x, y, z,
+    w)$-space. Now, we are stuck with an equation in four variables. With only
+    three degrees of freedom to work with when visualizing these objects we
+    encounter an obstacle. How do we know what these surfaces look like? This
+    brings us to the world of \emph{projective geometry}.
 
@@
 
-    The way we visualize these surfaces in four variables is by \emph{fixing}
-    one of the variables. This is called \emph{homogenizing} the polynomial
-    $h(x, y, z)$. We first want to introduce some definitions.
+    The way we work these surfaces in four variables is by introducing a new
+    variable making the polynomial terms all have the same degree.  This is
+    called \emph{homogenizing} the polynomial $h(x, y, z)$. We first want to
+    introduce some definitions.
 
@@
             \left( x_1, \ldots, x_{n+1} \right) \sim \left( x_1', \ldots,
-            \x_{n+1}' \right) \iff \left( x_1', \ldots, x_{n+1}' \right) =
+            x_{n+1}' \right) \iff \left( x_1', \ldots, x_{n+1}' \right) =
             \lambda \left( x_1, \ldots, x_{n+1} \right)
@@
 
+    \section{Algebraic Surfaces}
+    \label{sec:algebraic_surfaces}
+
+    We are now equipped with what we need to again look at our algebraic
+    surfaces defined by $h(x, y, z) = 0$. The solutions to this equation are
+    points in $\C^4$. We homogenize this polynomial by introducing the variable
+    $w$ and evaluating $h$ at the points $x \to x/w$, $y \to y/w$ and $z \to
+    z/w$ and then multiply by $w^d$ where $d$ is the degree of the polynomial
+    $h(x, y, z)$.
+
+    \begin{defn}[Homogenized polynomial]
+        Let $h(x, y, z)$ be a polynomial in $x, y, z$ of degree $d$. 
+        The homogenized polynomial $\hom(w, x, y, z)$ is given by
+        \begin{equation}
+            \notag
+            \hom(w, x, y, z) = w^d h \left( \tfrac{x}{w}, \tfrac{y}{w}, \tfrac{z}{w} \right).
+        \end{equation}
+        It follows that each non-zero term in this polynomial has the same
+        degree.
+    \end{defn}
+    The property that each non-zero term has the same degree in a homogeneous
+    polynomial gives rise to a very important property which relates directly
+    to the projective geometry introduced above.\footnote{This following
+    property can be seen directly by visualizing the homogenous polynomial
+    formed by a polynomial $g(x, y)$ and noticing that the figure does not
+    change when zooming in or out.}
+
+    \begin{crl}
+        If $\hom$ is an homogeneous polynomial formed by homogenizing the
+        polynomial $h(x, y, z)$ of degree $d$, then
+        \begin{equation}
+            \notag
+            \hom(\lambda w, \lambda x, \lambda y, \lambda z) = \lambda^d \hom(w, x, y, z).
+        \end{equation}
+    \end{crl}
+    \begin{proof}
+        We simply compute the left hand side:
+        \begin{align*}
+            \notag
+            \hom(\lambda w, \lambda x, \lambda y, \lambda z) &= (\lambda w)^d h
+            \left( \tfrac{\lambda x}{\lambda w}, \tfrac{\lambda x}{\lambda w},
+            \tfrac{\lambda x}{\lambda w}\right) \\ &= \lambda^d \hom(w, x, y,
+            z).
+        \end{align*}
+    \end{proof}
+    Since we are working with homogenous coordinates, the condition $\hom(w, x,
+    y, z) = 0$ is well defined on the equivalence class discussed above,
+    strictly due to this first corollary.
+
+    Now, if we can homogenize a polynomial, we can certainly dehomogenize it.
+    Given a homogeneous polynomial $\hom(w, x, y, z)$ we can dehomogenize the
+    polynomial by \emph{fixing} one of the variables.
+
+    \subsection{Initial examples of algebraic surfaces}
+    \label{sub:initial_examples_of_algebraic_surfaces}
+
+    Although mentioned before, we repeat the definition of an algebraic surface.
+    \begin{defn}[Algebraic surface]
+        A surface given by the equation $h(x, y, z) = 0$ is called \emph{algebraic}
+        if $h(x, y, z)$ is a polynomial in the variables $x, y, z$.
+        Mathematically,
+        \begin{equation}
+            \notag
+            h(x, y, z) = \sum^{}_{i, j, k} a_{ijk}x^iy^jz^k.
+        \end{equation}
+    \end{defn}
+
+    We are also interested in classifying the surfaces according to their
+    \emph{degree}. 
+    \begin{defn}[The degree of a surface]
+        Given a surface defined by the equation $h(x, y, z) = 0$ we define the
+        \emph{degree}, denoted $d$, of the surface to be the largest sum of
+        powers in the terms of the polynomial. Mathematically,
+        \begin{equation}
+            \notag
+            d = \max\{i + j + k\}
+        \end{equation}
+        where $i, j, k$ are the powers of $x, y$ and $z$ in the terms.
+    \end{defn}
+    Now, what do these surfaces actually look like? We give some examples that
+    is going to illustrate the above definition.
+
+    \begin{exmpl}
+        Let $h$ be a polynomial in $x, y, z$ given by
+        \begin{equation}
+            \notag h(x, y, z) = x^2 + y^2 - z^2. 
+        \end{equation}
+        We then look at the set of solutions to the equation $h(x, y, z) = 0$.
+        Plotting these yields a double-cone shaped figure. We see there is an
+        interesting quirk at the point $(0, 0, 0)$ where the surface seems to
+        pass through itself, formally we say that the surface exhibits a
+        \emph{singularity} at the point $(0, 0, 0)$, but we will come back to
+        that later.
+        \begin{figure}[h!]
+            \centering
+            \includegraphics[width=0.3\linewidth]{../pictures/double_cone.png}
+            \caption{The equation $h(x, y, z) = x^2 + y^2 - z^2 = 0$ gives rise
+            to a double-cone shaped algebraic surface.}
+            \label{fig:../pictures/double_cone}
+        \end{figure}
+        This surface has degree $d = 2$, and since all the terms have the same
+        degree the defining polynomial is homogenous, and hence the surface is
+        invariant under scaling.  
+    \end{exmpl}
+
+    While there is a lot of fun to be had with algebraic surfaces of degree $2$
+    there is a limit as to how exciting they are. Things get much more fun when
+    we look at surfaces of higher degree. Consider for example the following
+    surface of degree $3$.
+
+    \begin{exmpl}[The ding-dong cubic]
+        Let $h(x, y, z) = x^2 + y^2 - (1 - z)z^2$. Then the equation $h(x, y,
+        z) = 0$ defines a surface of degree $d = 3$ (attributed to the
+        $(1-z)z^2$ term). 
+        \begin{figure}[h!]
+            \centering
+            \includegraphics[width=0.3\linewidth]{../pictures/ding_dong.png}
+            \caption{The ding-dong cubic given by the equation $h(x, y, z) =
+            x^2 + y^2 - (1-z)z^2 = 0$.}
+            \label{fig:../pictures/ding_dong}
+        \end{figure}
+        This surface also exhibits a singularity at the point $(0, 0, 0)$ where
+        it passes through itself.
+    \end{exmpl}
+
+    Another very interesting  
+\clearpage
 \printbibliography
-
 \end{document}

### 2016-04-22 project/visualization/curves/curves.py

@@ -1,0 +1,21 @@
+import matplotlib.pyplot
+from numpy import arange
+from numpy import meshgrid
+
+
+matplotlib.pyplot.xkcd()
+
+delta = 0.025
+xrange = arange(-2.5, 5.0, delta)
+yrange = arange(-2.5, 2.5, delta)
+X, Y = meshgrid(xrange,yrange)
+
+# F is one side of the equation, G is the other
+F = Y**2
+G = X**2 + X**3 
+
+matplotlib.pyplot.axis('off')
+matplotlib.pyplot.contour(X, Y, (F - G), [0])
+
+matplotlib.pyplot.savefig('double_point.pdf')
+

### 2016-04-29 project/article_revised/src/article.tex
started revising the article

\documentclass{article}

\usepackage[]{amsthm} 
\usepackage[]{amsmath} 
\usepackage[]{amssymb}

\usepackage[]{graphicx} 
\usepackage[]{subcaption}

\theoremstyle{definition}
\newtheorem{definition}{Definition}
\newtheorem{corollary}{Corollary}
\newtheorem{example}{Example}

\usepackage[margin=2cm]{geometry}

\usepackage[]{mathpazo} 
\usepackage[]{palatino}

\usepackage[]{cleveref}

\newcommand{\R}{\mathbb{R}}
\newcommand{\A}{\mathbb{A}}
\newcommand{\V}{\mathbb{V}}
\renewcommand{\P}{\mathbb{P}}

\DeclareMathOperator{\rank}{rank}
\DeclareMathOperator{\col}{Col}
\title{Algebraic Surfaces}
\author{Ivar Haugal{\o}kken Stangeby}
\date{MAT2000 --- Spring 2016}

\begin{document}
\maketitle

\begin{abstract}
    In this paper we take a closer look at algebraic surfaces. With an emphasis
    on the visual aspects we look at singularities of surfaces, their
    symmetries, as well as techniques for resolving singularities. Unless
    specificed, the images are rendered in the software \textsc{Surfer}.
\end{abstract}

\tableofcontents

\section{Preliminaries}
\label{sec:preliminaries}

In algebraic geometry the objects of study are, amongst other, the set of
solutions to polynomial equations. Finding the zeroes of a quadratic equation
\begin{equation}
    \notag
    ax^2 + bx + c = 0
\end{equation} 
is something most people who have been exposed to elementary mathematics have
done. Later one learns how to find the zeroes of more general functions $f \in
K[x_1, x_2, \ldots, x_n]$ of either one or several variables, and you often
find functions that have no zeroes at all.

One of the central objects in algebraic geometry is the \emph{affine algebraic
variety} and the \emph{projective variety}. These are defined as the set of
common solutions to the zero-equations of a set of polynomials in some
\emph{polynomial ring}.

\begin{definition}[Polynomial ring]
    A \emph{polynomial ring} in $n$ variables $K[x_1, x_2, \ldots, x_{n}]$ over a
    field $K$ is the set of all functions on the form
    \begin{equation}
        \notag
        f(x) = \sum_{\alpha} p_\alpha \prod_{i=1}^{n} x_n^{\alpha_i}
    \end{equation}
    where $\alpha = (\alpha_1, \ldots, \alpha_n)$ is a multi-index and
    $p_\alpha = p_{\alpha_1\ldots\alpha_n}$ is a coefficient in $K$.
\end{definition}
An example of a polynomial ring is $\R[x, y]$. Elements include $f_1(x, y) =
x^2y + y + 1$ and $f_2(x, y) = \pi$. Given a set of polynomials $\left\{f_1, f_2, \ldots,
f_m\right\}$ in a polynomial ring $K \left[ x_1, x_2, \ldots, x_n \right]$ we can look at
their common zeroes. Specifically for polynomials in \emph{affine space} we
make the following definition.

\begin{definition}[Affine algebraic variety]
    Given a set of polynomials $\left\{ f_1, f_2, \ldots, f_m \right\}$ from a ring
    of polynomials, the set of points $(x_1, x_2, \ldots, x_n)$ in the affine space
    $\A^n$ that satisfy
    \begin{equation}
        \notag
        f_i(x_1, \ldots, x_n) = 0 \quad \text{for} \quad i = 1, \ldots, m
    \end{equation}
    is called an \emph{affine algebraic variety} and is denoted $\V(f_1, f_2,
    \ldots, f_m)$.
\end{definition}

Given the polynomial $f(x) = x^2 - 1$, the affine algebraic variety $\V(f)$ is
then the set of points $\left\{ 1, -1 \right\}$ since $f(-1) = f(1) = 0$.
Given an additional polynomial $g(x) = x-1$ we have $\V(f, g) = \left\{ 1
\right\}$. Finally, if we also consider the polynomial $h(x) = x^2$, then
$\V(f, g, h) = \emptyset$. We therefore see that the variety of a set of
function is an intersection of the varieties of the individual functions.

In this paper we will also deal a lot with singularities of our algebraic
surfaces. We make a formal definition.
\begin{definition}[Singularity]
    A singularity of the variety $V = \V(f_1, f_2, \ldots, f_m)$ is a point $p$ such
    that the Jacobian matrix
    \begin{equation}
        \notag
        \begin{bmatrix}
            \frac{\partial f_1}{\partial x_1} && \hdots && \frac{\partial f_1}{\partial x_n} \\
            \vdots && \ddots && \vdots \\
            \frac{\partial f_m}{\partial x_1} && \hdots && \frac{\partial f_m}{\partial x_n}
        \end{bmatrix}(p)
    \end{equation} has rank strictly less than $\min(m, n)$.
\end{definition}

A very useful consequence of this definition deal with varieties of single
polynomials.
\begin{corollary}
    \label{crl:single_poly}
    A point $p$ is a singularity of the variety $\V(f)$ if and only if the
    partial derivatives of $f$ vanish identically at $p$. Mathematically this
    is expressed as ${\partial f(p)}/{\partial x_i} = 0$ for $i = 1, 2, \ldots,
    n$.
\end{corollary}
\begin{proof}
    We recall that the \emph{rank} of a matrix $A$ is the dimension of its
    column space. Note that $\min(m, n)$ is in this case equal to $1$ since we
    have $n$ variables but only one function.
    For the Jacobian matrix
    \begin{equation}
        \notag
        A = \begin{bmatrix}
            \frac{\partial f_1}{\partial x_1} && \hdots && \frac{\partial f_1}{\partial x_n}
        \end{bmatrix}(p)
    \end{equation} 
    to have rank strictly less then one, its column space has to have dimension
    zero. In linear algebra terms, this simply means that the matrix have no
    pivot columns, and hence each column must be zero. Therefore, for $p$ to be
    a singularity of $\V(f)$ all its partial derivatives must vanish
    identically.
\end{proof}

\section{Algebraic Surfaces}
\label{sec:algebraic_surfaces}

In this section we try to motivate the definition of an algebraic surface and
look at some examples, and hopefully give an intuitive explanation of what
exactly we are looking at. Looking at real algebraic curves is a good entry
point.

\subsection{An Informal Introduction}
\label{sub:an_informal_introduction}

From elementary mathematics one learns about real valued functions, $f(x)$, and
how to graph these functions by setting $y = f(x)$ and plotting points in the
$(x, y)$-plane. Now, the \emph{graph of a function} is something of a
pecuiliarity, because it comes with some restrictions. Not all curves in the
$(x, y)$-plane correspond to a specific function. The method for verifying
whether a certain ''graph'' correspond to a function or not, typically taught
in school, is the \emph{vertical line test}.

Having an equation $y = f(x)$ we can form what we call the \emph{equation of a
curve at zero}. We define a new function in two variables
\begin{equation}
    \notag
    g(x, y) = y - f(x) = 0.
\end{equation}
If the function $f(x)$ is a polynomial in the variable $x$ we call the function
graph of $f(x)$ \emph{algebraic}. If the distinction between a graph and a
curve is to be justified there has to be some curves that are not graphs. If we
look at what happens when you set $y^2 = x^3$, then the plotted curve does not
pass the vertical line test, and is therefore something different from a
function graph. Similarily, the equation $y^2 = x^3 - x^2$ defines a curve that
is again not a graph. These curves also happen to exhibit \emph{singular
points} at $(0, 0)$.

With what we have so far, we can move up a dimension. Instead of considering
curves in the $(x, y)$-plane we can look at surfaces in $(x, y, z)$-space.
Following the same procedure as above, we define a new function $h(x, y, z) = z
- g(x, y)$. The surfaces are given by equations on the form $h(x, y, z) = 0$.
These surfaces are \emph{algebraic} if $h(x, y, z)$ is a polynomial in the
variables $x, y, z$.

\begin{example}
    \label{exmpl:algebraic_surfaces}
    The following equations satisfy the definition of an algebraic surface
    \begin{enumerate}
        \item $h_1(x, y, z) = z - xy = 0$,
        \item $h_2(x, y, z) = z - x^2 - y^2 = 0$,
        \item $h_3(x, y, z) = z - (2x^2-y)(y-x^2)$,
    \end{enumerate}
    where as the following does not satisfy the definition:
    \begin{enumerate}
        \item $g_1(x, y, z) = \sin(xy) + z$,
        \item $g_2(x, y, z) = \tan^2(x) + \pi zy$.
    \end{enumerate}
    The examples of algebraic surfaces are shown visualized in
    \cref{fig:algebraic_surfaces}.

    \begin{figure}[]
        \centering
        \begin{subfigure}[t]{0.3\textwidth}
            \includegraphics[height=0.6\textwidth, width=0.6\textwidth]{../pictures/example_one.png}
            \caption{$h_1(x, y, z) = 0$.}
        \end{subfigure}
        ~
        \begin{subfigure}[t]{0.3\textwidth}
            \includegraphics[height=0.6\textwidth, width=0.6\textwidth]{../pictures/example_two.png}
            \caption{$h_2(x, y, z) = 0$.}
        \end{subfigure}
        ~
        \begin{subfigure}[t]{0.3\textwidth}
            \includegraphics[height=0.6\textwidth, width=0.6\textwidth]{../pictures/example_three.png}
            \caption{$h_3(x, y, z) = 0$.}
        \end{subfigure}
        \caption{The algebraic surfaces from \cref{exmpl:algebraic_surfaces}.}
        \label{fig:algebraic_surfaces}
    \end{figure}
\end{example}

We complete this section by formally defining an algebraic surface.
\begin{definition}[Algebraic surface]
    An \emph{algebraic surface} is an algebraic variety of dimension
    two.\footnote{The dimension of an algebraic variety is a bit hard to grasp,
    and since we only will deal with algebraic surfaces in this article we do
    not spend any time classifying the different kinds of varieties.}

    We define the \emph{degree} of the algebraic surface to be highest combined
    power in the defining polynomials terms, and denote this $d$, i.e.,
    \begin{equation}
        \notag
        d = \max_{\alpha}\left\{\alpha_1 + \alpha_2 + \ldots + \alpha_n\right\}
    \end{equation}
    where the $\alpha_i$ is the power of $x_i$ in a term.
\end{definition}

\section{Singularities}
\label{sec:singularities}

The singularities of an algebraic surface is one of the more important elements
that give the surfaces its visual appeal. I am sure many people would agree
with me that surfaces with many singularities, and with singularities in
interesting configurations, are more exciting to look at then the surfaces that
show no singularities.

We first take a look at the number of singularities that various surfaces have.

\subsection{The Number of Singularities}
\label{sub:the_number_of_singularities}

For an algebraic surface of degree $d$ we let $\mu(d)$ denote the maximum
number of singularities such a surface \emph{can} exhibit. We first look at
some examples that help illustrate this.

\begin{example}[The plane has no singularities]
    The defining polynomial of a plane in three dimensions is 
    \begin{equation}
        \notag
        f(x, y, z) = ax + by + cz + d.
    \end{equation}
    The corresponding surface is then $V = \V(f)$. We want to find the
    singularities of $V$. Applying \cref{crl:single_poly} we only need consider
    the partial derivatives and find the points in $V$ where they vanish. We
    solve the equations
    \begin{align*}
        \frac{\partial f}{\partial x} = \frac{\partial f}{\partial y} = \frac{\partial f}{\partial z} = 0.
    \end{align*} 
    This gives $a = 0$, $b = 0$ and $c = 0$ which again forces $d = 0$ since
    otherwise our point $(x, y, z)$ would not be a zero of $f$. This however,
    tells us that $f(x, y, z) = 0$ which does not define any surface. Hence,
    the plane has no singularities. The plane is an algebraic surface of degree
    $d = 1$, so consequently $\mu(1) = 0$.
\end{example}

\begin{example}[The Cayley cubic]
    The Cayley cubic is defined by the polynomial
    \begin{equation}
        \notag
        f(x, y, z) = xyz + (1 - x - y - z)(yz + xz + xy).
    \end{equation}
    Again, we are interested in the singularities of $V = \V(f)$. Once again,
    we apply \cref{crl:single_poly} and look at the partial derivatives and
    find their zeroes. This gives the four points $(0, 0, 0)$, $(0, 0, 1)$,
    $(0, 1, 0)$ and $(1, 0, 0)$.

    It can also be shown that there cannot be \emph{more} than four
    singularities for a third degree surface. Hence $\mu(3) = 4$.
\end{example}

\subsection{Symmetric singularities}
\label{sub:symmetric_singularities}

\section{Resolving Singularities}
\label{sec:resolving_singularities}

We call a variety that exhibits no singularities \emph{smooth}. It is often
possible to study a smooth copy of a variety with singularities if we move from
affine space to projective space.

\subsection{Projective Space}
\label{sub:projective_space}

In this section we define the notion of a \emph{projective space}.

\subsection{Blowups}
\label{sub:blowups}

Informally, when we blow up a point $p$ in $\A^n$, we replace the point $p$ by
an entire copy of a projective space $\P^{n-1}$ and try to leave the rest of
the space $\A^n$ unchanged. We will give a few examples of this.

\begin{definition}[Blowup of $\A^n$ at $p$]
   Let the blowup-surface $B$ be the set of all pairs consisting of a point $p
   \in \A^n$ and a line from $p$ through the origin.  Then $B \subseteq
   \A^n\times\P^{n-1}$. We write
   \begin{align*}
       \notag
       B &= \left\{ (p, L) \mid p \in \A^n, p \in L \right\} \\
         &= \left\{ (x_1, \ldots, x_n ; y_1: y_2: \ldots :y_n) \mid x_iy_j =
           x_jy_i \quad 0 \leq i < j \leq n \right\}.
   \end{align*}
\end{definition}

In order to aid us with this, we have the following result:
\begin{corollary}
    \label{crl:minors}
    Given a point $x = (x_1, \ldots, x_n)$ in the space $\A^n$ and a point $y =
    \left( y_1 : \ldots : y_n \right)$ in the space $\P^{n-1}$ the point
    $x$ is a multiple of $y$ if and only if the matrix
    \begin{equation}
        \notag
        A = 
        \begin{bmatrix}
            x_1 & x_2 & \hdots & x_n \\
            y_1 & y_2 & \hdots & y_n
        \end{bmatrix}
    \end{equation}
    has rank less than or equal to one. Similarly, the rank is less than or
    equal to one only if each $2\times2$-minor $x_iy_j - x_jy_i = 0$.\footnote{Figure this out properly.}
\end{corollary}
\begin{proof}
    In order to verify this, we recall that $\rank(A) = \rank(A^T)$, so we
    chose to work with the transposed matrix instead. Assume that $x$ is a
    multiple of $y$, so $y = \lambda x$ for some scalar $\lambda$.  Then
    \begin{align*}
        A^T = \begin{bmatrix}
            x_1 & \lambda x_1 \\
            \vdots & \vdots \\
            x_n & \lambda x_n
        \end{bmatrix}.
    \end{align*}
    Assume that the point $x$ is non-zero. Then we have exactly one pivot
    column in this matrix, hence $\dim(\col(A^T)) = 1$. If $x = 0$, then we have
    no pivot columns, so $\dim(\col(A^T)) = 0$. In any case, $\rank(A) =
    \rank(A^T) \leq 1$.

    For the second claim, since $A^T$ has rank 1, there cannot exists any
    non-zero minors $r\times r$ for $r > 1$. Hence all $2\times 2$ minors must
    vanish.
\end{proof}

\begin{example}[The Node]
    The defining polynomial of the node is given by
    \begin{equation}
        \label{eq:node}
        f(x, y) = y^2 - x^2 - x^3,
    \end{equation}
    so the curve is $C = \V(f)$. This curve lives in the affine plane $\A^2$.
    The blow-up $B_p$ of $C$ at the point $p = (0, 0)$ lives in $\A^2 \times
    \P^1$. For a point $(x, y) \in C$ we figure out what line it passes through
    by applying \cref{crl:minors} with homogeneous coordinates $(r, s)$.
    \begin{equation}
        \notag
        \begin{bmatrix}
            x & y \\
            r & s
        \end{bmatrix}
    \end{equation}
    has rank less than or equal to one when $xs - yr = 0$ (our only
    $2\times2$-minor in this case). This gives $xs = yr$. Since $r$ and $s$ are
    homogeneous coordinates, we have $(r : s) = (1 : s)$ and $(r : s) = (r :
    1)$. We can therefore eliminate either $r$ or $s$ in our equation.

    \begin{description}
        \item[Case 1] Setting $s = 1$ we get $x = yr$. Eliminating $x$ from
            \cref{eq:node} gives a new equation $g(y, r) = y^2 - y^2r^2 -
            y^3r^3 = y^2(1 - r^2 - yr^3)$. This can be visualized in the $(y,
            r)$-plane.
        \item[Case 2] Setting $r = 1$ we get $y = xs$. We can now eliminate $y$
            from \cref{eq:node} which gives another new equation $h(x, s) =
            x^2s^2 - x^2 - x^3 = x^2(s^2 - 1 - x)$ which can be visualized in
            the $(x, s)$-plane.
    \end{description}
\end{example}

\end{document}

### 2016-04-29 project/article_revised/src/aux/prop_7.tex
started revising the article

\documentclass{amsart}

\begin{document}
    Let $(x_1, \ldots, x_n)$ be a multiple of $(y_1, \ldots, y_n)$.  We can
    then write the matrix
    \begin{equation}
        \notag
        \begin{bmatrix}
            x_1 && \hdots && x_n \\
            y_1 && \hdots && y_n
        \end{bmatrix} = 
        \begin{bmatrix}
            x_1 && \hdots && x_n \\
            \lambda x_1 && \hdots && \lambda x_n \\
        \end{bmatrix}
    \end{equation}
    for some scalar $\lambda$. However, we can now subtract a multiple
    $\lambda$ of the first row from the second row and achieve
    \begin{equation}
        \notag
        \begin{bmatrix}
            x_1 && \hdots && x_n \\
            0 && \hdots && 0
        \end{bmatrix}.
    \end{equation} If we now let $x_i$ be the smallest non-zero component for
    $1 \leq i \leq n$ and divide the first row by $x_i$ we achieve the
    following matrix
    \begin{equation}
        \notag
        \begin{bmatrix}
            0 && \hdots && 1 && \frac{x_i+1}{x_i} && \hdots && \frac{x_n}{x_i} \\
            0 && \hdots && 0 && 0 && \hdots && 0
        \end{bmatrix}.
    \end{equation}
    This matrix has exactly one pivot-column if there is at least one $i$ for
    which $x_i \neq 0$ and exactly zero pivot columns if $x_i = 0$ for all $i$.
    Consequently, the rank of this matrix is either 1 or 0, and more
    specifically, less than or equal to 1.
\end{document}

### 2016-04-29 project/visualization/curves/curves.py
node blowup done

@@ -6,16 +6,18 @@ from numpy import meshgrid
 matplotlib.pyplot.xkcd()
 
 delta = 0.025
-xrange = arange(-2.5, 5.0, delta)
+xrange = arange(-5, 5.0, delta)
 yrange = arange(-2.5, 2.5, delta)
 X, Y = meshgrid(xrange,yrange)
 
 # F is one side of the equation, G is the other
-F = Y**2
-G = X**2 + X**3 
+# F = X**2 * (Y**2 - 1 - X)
+F = X**2*(1 - Y**2 - X*Y**3)
+G = 0
 
-matplotlib.pyplot.axis('off')
+matplotlib.pyplot.axis('on')
+matplotlib.pyplot.grid('on')
 matplotlib.pyplot.contour(X, Y, (F - G), [0])
 
-matplotlib.pyplot.savefig('double_point.pdf')
+matplotlib.pyplot.savefig('case_1.pdf')
 

### 2016-05-20 project/article_revised/src/article.tex
node blowup done
backup

@@
 
+% {{{ Preamble 
 \usepackage[]{amsthm} 
@@
 \usepackage[]{amssymb} 
+\usepackage[]{thmtools}
+
+\usepackage[]{fontspec}
+
+% Code listings
+\usepackage[]{listings}
+\usepackage[usenames, dvipsnames, svgnames, table]{xcolor}
+\usepackage[]{mdframed}
+\usepackage[]{calc}
+
+\definecolor{TMcodeBackground}{RGB}{224, 224, 224}
+\definecolor{TMcodeFrame}{RGB}{109, 108, 109}
+\definecolor{TMgreen}{RGB}{14, 191, 48}
+\definecolor{TMorange}{RGB}{243, 126, 25}
+\definecolor{TMred}{RGB}{230, 6, 85}
+
+\defaultfontfeatures{Ligatures=TeX}
+\setmonofont{Source Code Pro Light}[BoldFont=Source Code Pro]
+
+\mdfdefinestyle{TMstyleCode}{
+            skipabove=4mm,
+            skipbelow=0mm,
+            %remove borders
+            rightline=false,
+            topline=false,
+            bottomline=false,
+            linewidth=1mm,
+            %margins
+            innertopmargin=2mm,
+            innerleftmargin=0mm,
+            innerbottommargin=0mm,
+            innerrightmargin=10pt,
+            linecolor=TMcodeFrame,
+            backgroundcolor=TMcodeBackground
+}
+
+\lstdefinestyle{TMstyle}{
+    showstringspaces=false,
+    numbers=left,
+    numbersep=7mm,
+    numberstyle=\color{Black},
+    stepnumber=1,
+    tabsize=3,
+    breakatwhitespace=false,
+    breaklines=true,
+    captionpos=b,
+    basicstyle=\color{Black}\ttfamily,
+    commentstyle=\color{TMgreen}\ttfamily,
+    keywordstyle=\color{TMorange}\bfseries,
+    stringstyle=\color{TMred},
+    frame=leftline,
+    framesep=0mm,
+    xleftmargin=3mm,
+    framesep=2mm,
+    framerule=0mm,
+    abovecaptionskip=5mm,
+    aboveskip=\baselineskip,
+    belowskip=\baselineskip
+}
+
+\lstnewenvironment{TMcode}[3]
+{
+    \lstset{style=TMstyle, language=#1, caption=#2}
+    \mdfsetup{style=TMstyleCode}
+    \mdframed
+    \hspace*{3mm}
+    \minipage{0.75cm}
+    \includegraphics[width=\linewidth]{../pictures/code2.png}
+    \endminipage
+    \hspace*{1mm}
+    \minipage{\textwidth-1.05cm}
+        {\sffamily\Large #3}
+    \endminipage
+    \vspace*{-2mm}
+}
+{
+    \endmdframed
+}
 
@@
 
-\theoremstyle{definition}
-\newtheorem{definition}{Definition}
-\newtheorem{corollary}{Corollary}
-\newtheorem{example}{Example}
-
+\declaretheorem[style=definition,qed=$\clubsuit$]{definition}
+\declaretheorem[style=definition,qed=$\spadesuit$,sibling=definition]{example}
+\declaretheorem[style=definition,qed=$\diamondsuit$]{corollary}
 \usepackage[margin=2cm]{geometry}
 
-\usepackage[]{mathpazo} 
-\usepackage[]{palatino}
-
+\usepackage[]{url} 
+\urlstyle{sf}
+\usepackage[]{makeidx} 
+\makeindex
 \usepackage[]{cleveref}
@@
 \newcommand{\R}{\mathbb{R}}
+\newcommand{\C}{\mathbb{C}}
 \newcommand{\A}{\mathbb{A}}
 \newcommand{\V}{\mathbb{V}}
+\newcommand{\Bu}[2]{\mathrm{Bl}_{#1}\A^{#2}}
 \renewcommand{\P}{\mathbb{P}}
@@
 \DeclareMathOperator{\col}{Col}
+% }}}
+
 \title{Algebraic Surfaces}
@@
 One of the central objects in algebraic geometry is the \emph{affine algebraic
-variety} and the \emph{projective variety}. These are defined as the set of
+variety}\index{variety!affine algebraic variety} and the \emph{projective
+variety}\index{variety!projective variety}. These are defined as the set of
 common solutions to the zero-equations of a set of polynomials in some
-\emph{polynomial ring}.
+\emph{polynomial ring}\index{polynomial ring}.
 
@@
 f_m\right\}$ in a polynomial ring $K \left[ x_1, x_2, \ldots, x_n \right]$ we can look at
-their common zeroes. Specifically for polynomials in \emph{affine space} we
+their common zeroes. Specifically for polynomials in \emph{affine space}\index{space!affine} we
 make the following definition.
@@
 \begin{proof}
-    We recall that the \emph{rank} of a matrix $A$ is the dimension of its
+    We recall that the \emph{rank}\index{rank} of a matrix $A$ is the dimension of its
     column space. Note that $\min(m, n)$ is in this case equal to $1$ since we
@@
 how to graph these functions by setting $y = f(x)$ and plotting points in the
-$(x, y)$-plane. Now, the \emph{graph of a function} is something of a
+$(x, y)$-plane. Now, the \emph{graph of a function}\index{graph} is something of a
 pecuiliarity, because it comes with some restrictions. Not all curves in the
@@
 Having an equation $y = f(x)$ we can form what we call the \emph{equation of a
-curve at zero}. We define a new function in two variables
+curve at zero}\index{equation at zero}. We define a new function in two variables
 \begin{equation}
@@
 is again not a graph. These curves also happen to exhibit \emph{singular
-points} at $(0, 0)$.
+points}\index{singular point} at $(0, 0)$.
 
@@
 - g(x, y)$. The surfaces are given by equations on the form $h(x, y, z) = 0$.
-These surfaces are \emph{algebraic} if $h(x, y, z)$ is a polynomial in the
-variables $x, y, z$.
+These surfaces are \emph{algebraic}\index{surface!algebraic surface} if $h(x,
+y, z)$ is a polynomial in the variables $x, y, z$.
 
@@
 \begin{definition}[Algebraic surface]
-    An \emph{algebraic surface} is an algebraic variety of dimension
-    two.\footnote{The dimension of an algebraic variety is a bit hard to grasp,
-    and since we only will deal with algebraic surfaces in this article we do
-    not spend any time classifying the different kinds of varieties.}
+    An \emph{algebraic surface}\index{surface!algebraic surface} is an
+    algebraic variety of dimension two.\footnote{The dimension of an algebraic
+    variety is a bit hard to grasp, and since we only will deal with
+    algebraic surfaces in this article we do not spend any time classifying
+    the different kinds of varieties.}
 
-    We define the \emph{degree} of the algebraic surface to be highest combined
-    power in the defining polynomials terms, and denote this $d$, i.e.,
+    We define the \emph{degree}\index{surface!degree} of the algebraic surface
+    to be highest combined power in the defining polynomials terms, and denote
+    this $d$, i.e.,
     \begin{equation}
@@
 
-\subsection{Symmetric singularities}
+\subsection{Symmetric Singularities}
 \label{sub:symmetric_singularities}
@@
 
-We call a variety that exhibits no singularities \emph{smooth}. It is often
-possible to study a smooth copy of a variety with singularities if we move from
-affine space to projective space.
+We call a variety that exhibits no singularities \emph{smooth}\index{smooth}.
+It is often possible to study a smooth copy of a variety with singularities if
+we move from affine space to projective space.
 
@@
 
-In this section we define the notion of a \emph{projective space}.
+In this section we define the notion of a \emph{projective
+space}\index{space!projective}.  Informally, projective space formalizes the
+notion of parallel lines intersecting at infinity.
 
+The general definition of projective $n$ space is the set of lines in
+$\C^{n+1}$ passing through the origin.
+\begin{definition}[The projective space of dimension $n$]
+    Let $\sim$ be an equivalence relation on $\C^{n+1} \setminus 0$ defined by
+    \begin{equation}
+        \notag
+        \left( x_1, \ldots, x_{n+1} \right) \sim \left( x_1', \ldots,
+        x_{n+1}' \right) \iff \left( x_1', \ldots, x_{n+1}' \right) =
+        \lambda \left( x_1, \ldots, x_{n+1} \right)
+    \end{equation}
+    where $\lambda$ is some non-zero complex number. We denote the equivalence
+    class of $\left( x_1, \ldots, x_{n+1} \right)$ by $\left[ x_1 : \ldots :
+    x_{n+1} \right]$ and define the \emph{projective space of dimension $n$} as
+    the set
+    \begin{equation}
+        \notag
+        \P^n{\C} = \left\{ \left[ x_1 : \ldots : x_{n+1} \right] \mid
+        \left( x_1, \ldots, x_{n+1} \right) \in \C^{n+1} \setminus
+        0 \right\}.
+    \end{equation}
+\end{definition}
 \subsection{Blowups}
@@
 
-\begin{definition}[Blowup of $\A^n$ at $p$]
-   Let the blowup-surface $B$ be the set of all pairs consisting of a point $p
-   \in \A^n$ and a line from $p$ through the origin.  Then $B \subseteq
+\begin{definition}[Blowup of $\A^n$ at $p$]\index{blowup!point}
+    \label{defn:blowup_point}
+    We let the blowup-surface\index{surface!blowup} $\Bu{p}{n}$ be the set of all
+    pairs consisting of a point $p
+    \in \A^n$ and a line from $p$ through the origin.  Then $\Bu{p}{n} \subseteq
    \A^n\times\P^{n-1}$. We write
@@
        \notag
-       B &= \left\{ (p, L) \mid p \in \A^n, p \in L \right\} \\
+       \Bu{p}{n} &= \left\{ (p, L) \mid p \in \A^n, p \in L \right\} \\
          &= \left\{ (x_1, \ldots, x_n ; y_1: y_2: \ldots :y_n) \mid x_iy_j =
@@
     has rank less than or equal to one. Similarly, the rank is less than or
-    equal to one only if each $2\times2$-minor $x_iy_j - x_jy_i = 0$.\footnote{Figure this out properly.}
+    equal to one only if each $2\times2$-minor $x_iy_j - x_jy_i =
+    0$.\footnote{Figure this out properly.}
 \end{corollary}
@@
     \end{description}
+    \begin{figure}[h!]
+        \centering
+        \begin{subfigure}[t]{0.3\textwidth}
+            \includegraphics[width=\textwidth]{../pictures/case_1.pdf} 
+            \caption{The function $g$ visualized in the $(y, r)$-plane.}
+        \end{subfigure}
+        ~
+        \begin{subfigure}[t]{0.3\textwidth}
+            \includegraphics[width=\textwidth]{../pictures/case_2.pdf} 
+            \caption{The function $h$ visualized in the $(x, s)$-plane.}
+        \end{subfigure}
+        \caption{The affine charts of the blow-up. Note that each of the curves
+        intersect the line vertical axis twice. This corresponds to the two
+        singularities of the original curve.}
+    \end{figure}
+    Note that the functions $g$ and $h$ are homogeneous functions, in the sense
+    that $x$ and $y$ are regarded as constant and are therefore of degree
+    zero.\footnote{Figure out why.} It is now of interest to see how these
+    \emph{affine charts} relate to the original curve.
 \end{example}
 
+Sometimes it is not enough to only blow up the surface at a single point $p$.
+We may instead need to blow up the surface at a singular line, or a singular
+curve, or in general, any other set. We make an anologous definition to
+\cref{defn:blowup_point} however this time for the blowup of a line $L$ and
+for our purposes we specialize it to $\A^3$.
+
+\begin{definition}[Blowup of $\A^3$ along a line $L$]\index{blowup!line}
+    We let the blowup-surface $\Bu{L}{3}$ be the set of all pairs consisting of
+    a point $p \in \A^3$ and a point $\lambda \in \P^{1}$ such that
+    \begin{equation}
+        \notag
+        \Bu{L}{3} = \left\{ (p, \lambda) \in \A^3\times\P^1 \mid P \in H_\lambda \right\}
+    \end{equation}
+    where $H_\lambda$ is the unique plane passing through $L$ and
+    $P$.
+    \footnote{I think of $\lambda$ as a representative of the distinct lines in
+        $\A^2$, this can therefore be thought of as a point in $\P^1$.  Each
+        line passing through the origin in $\A^2$ (represented by $\lambda$)
+        paired with the line $L$ uniquely determines a plane $H_\lambda$.}
+\end{definition}
+
+With this definition, we tackle another example of a blow-up, where it does not
+suffice to only blow up the singularity in the origin.
+
+\begin{example}[Blowing up a line]
+    We want to study the variety $\V(x^2y - z^2)$. Not only has this surface a
+    singularity in the origin, but it also has a singularity for each point
+    along the line $x = z = 0$. It is visualized in \cref{fig:blowup_line}
+    where the singularities are easily pointed out.
+
+    \paragraph{Attempt one:}
+    \label{par:attempt_one}
+
+    If we were to proceed by trying to only blow up the origin, we quickly
+    encounter a problem.  Following the standard procedure, we get the defining
+    equations $xs = yr$, $xt = zr$ and $yt = zs$ with $r, s$ and $t$ are
+    homogeneous coordinates.
+
+    We consider the first affine chart by setting $s = 1$. We can then
+    eliminate $x$ and $y$ from $x^2y - z^2 = 0$. This yields the equation
+    \begin{equation}
+        \notag
+        y^2(r^2y - t^2) = 0.
+    \end{equation}
+    However, if we assume that $y \neq 0$ we can divide out the
+    \emph{exceptional divisor}\index{exceptional divisor} $y^2$ which results
+    in the equation $r^2y - t^2 = 0$. This however gives us a problem, because
+    this equation describes the same surface as our original surface, we have
+    therefore not resolved any singularities at all. So, clearly we have to try
+    a different approach. We try to blow up the entire line of singularity at
+    once, using the former definition.
+    \begin{figure}[]
+        \centering
+        \includegraphics[width=0.3\linewidth]{../pictures/blowup_line.png}
+        \caption{The surface defined by $\left\{ x^2y - z^2 = 0\right\}$. This
+        has a line of singularities along the $y$-axis, for $x = z = 0$.}
+        \label{fig:blowup_line}
+    \end{figure}
+
+    \paragraph{Attempt two:}
+    \label{par:attempt_two}
+
+    We need to find an equation for the plane $H_\lambda$. We want both the
+    $y$-axis and the point $(x, y, z)$ to lie in the plane $rx + sy + tz = 0$.
+    Chosing the vector $(0, 1, 0)$ as a representative for the $y$-axis we compute the normal vector.
+    \begin{equation}
+        \notag
+        n = \begin{vmatrix}
+            \hat{i} & \hat{j} & \hat{k} \\
+            r & s & t \\
+            0 & 1 & 0
+        \end{vmatrix} = (-t, 0, r).
+    \end{equation}
+    The equation for the plane is now given by
+    \begin{equation}
+        \notag
+        H_\lambda : -tx + 0y + rz = rz - tx = 0.
+    \end{equation}
+    We can now consider the two affine charts obtained by setting $r = 1$ and $t = 1$ respectively.
+    \paragraph{Case 1:} 
+    \label{par:case_1_}
+    Setting $r = 1$ yields $z = tx$ which when plugged into $x^2y - z^2 = 0$
+    gives us $x^2y - (t^2x^2) = x^2(y - t^2) = 0$. This affine chart can be
+    visualized in the $(x, y, t)$-plane.
+
+    \paragraph{Case 2:}
+    \label{par:case_2_}
+    Setting $t = 1$ yields $x = rz$ which when plugged into $x^2y-z^2 = 0$
+    gives the equation $r^2z^2y - z^2 = r^2(z^2y - 1) = 0$. This can be
+    visualized in the $(r, y, z)$-plane.
+
+    \begin{figure}[h!]
+        \centering
+        \begin{subfigure}[t]{0.3\textwidth}
+            \includegraphics[width=\textwidth]{../pictures/line_blowup_affine_1.png}
+            \caption{The surface given by the equation $x^2(y - t^2) = 0$. The
+            curve of intersection between this surface and the exceptional
+            divisor looks like it can be ''pinched'' into a double line, i.e., the
+            original line of singularity.}
+        \end{subfigure}
+        ~
+        \begin{subfigure}[t]{0.3\textwidth}
+            \includegraphics[width=\textwidth]{../pictures/line_blowup_affine_2.png} 
+            \caption{The surface given by the equation $r^2(z^2y - 1) = 0$. The
+            intersection between this surface and the exceptional divisor can
+        also be ''fused'' together to form a double line, by laying the
+    surfaces flat on top of eachother.}
+        \end{subfigure}
+        \\
+        \begin{subfigure}[t]{0.3\textwidth}
+            \includegraphics[width=\textwidth]{../pictures/affine_1_exceptional.pdf} 
+        \end{subfigure}
+        ~
+        \begin{subfigure}[t]{0.3\textwidth}
+            \includegraphics[width=\textwidth]{../pictures/affine_2_exceptional.pdf} 
+        \end{subfigure}
+        \caption{The two affine charts of the variety $\V(x^2y-z^2)$. The top
+            row is rendered by \textsc{Surfer}, and the bottom row rendered by
+            \textsc{Mathematica.} The exceptional divisor in the top row has
+            been divided out and is therefore not shown because \textsc{Surfer}
+        does a poor job at rendering planes. As seen, the charts are smooth and
+    exhibits no singularities. The intersection with the exceptional divisor is
+rendered in blue in the bottom row.}
+    \end{figure}
+\end{example}
+
+While in the above example we blew up a very specific line of singularities,
+namely the $y$-axis, it can be of interest to talk a little bit about how to
+blow up general lines. It all boils down to generalizing the construction of
+the plane for a more general line $L$. Assuming $L$ is parametrized by $x = at,
+y = bt, z = ct$, with $a, b$ and $c$ non-zero, we can form the symmetric
+equations 
+\begin{equation}
+    \notag
+    t = \frac{x}{a} = \frac{y}{b} = \frac{z}{c}.
+\end{equation}
+The line $L$ is then completely determined by two of the three possible
+relations $xb - ya = 0$, $xc - za = 0$ and $yc - zb = 0$.  In order to find the
+respective plane, we are interested in knowing when the plane $rx + sy + tz =
+0$ can be expressed in terms of the line given by $xb - ya = 0$ and $yc - zb =
+0$. This can be solved as a matrix problem. Given the matrix
+\begin{equation}
+    \notag
+    A = \begin{bmatrix}
+        b & 0 & r \\
+        -a & c & s \\
+        0 & -b & t
+    \end{bmatrix}
+\end{equation}
+one of the columns of $A$ is linearly dependent of the two others if $\det(A) =
+0$. Solving $\det(A) = 0$ gives the equation
+\begin{equation}
+    \notag
+    \det(A) = b(ct + sb) + rab = 0.
+\end{equation}
+We can now solve for either one of $r, s$ and $t$. Solving for $r$ we get $r =
+-( ct + sb ) / a$ which we can substitute back into the equation for our plane.
+The plane is then determined by 
+\begin{equation}
+    \notag
+    \frac{-ct - sb}{a}x + sy + tz = 0.
+\end{equation}
+Multiplying through by $a$ we achieve
+\begin{align*}
+    (-sb - ct)x + sya + tza &= -sbx - ctx + sya + tza \\
+                            &= t(za - cx) + s(ya - bx) = 0.
+\end{align*}
+\clearpage
+\appendix
+
+\newgeometry{margin=4cm}
+
+\section{Code Samples}
+\label{sec:code_samples}
+
+\lstset{deletekeywords={xrange}}
+\begin{TMcode}{python}{curves.py - \url{http://stackoverflow.com/a/2484594}}{Plotting algebraic curves}
+import matplotlib.pyplot
+from numpy import arange
+from numpy import meshgrid
+
+delta = 0.025
+xrange = arange(-5, 5.0, delta)
+yrange = arange(-2.5, 2.5, delta)
+X, Y = meshgrid(xrange,yrange)
+
+# F is one side of the equation, G is the other
+F = X**2*(1 - Y**2 - X*Y**3)
+G = 0
+
+matplotlib.pyplot.axis('on')
+matplotlib.pyplot.grid('on')
+matplotlib.pyplot.contour(X, Y, (F - G), [0])
+
+matplotlib.pyplot.savefig('output.pdf')
+\end{TMcode}
+
+\clearpage
+\begin{TMcode}{mathematica}{intersection.nb - \url{http://community.wolfram.com/groups/-/m/t/178218?p_p_auth=lT7er3Rd}}{Plotting the intersection between to surfaces}
+
+(* affine chart and exceptional divisor *)
+h = z^2 y - 1;
+g = x^2;
+
+ContourPlot3D[
+    {h == 0, g == 0}, 
+    (* bounding box dimensions *)
+    {x, -3, 3}, 
+    {y, -3, 3}, 
+    {z, -3, 3}, 
+    MeshFunctions -> {
+        Function[ {x, y, z, f}, h - g ]
+    }, 
+    MeshStyle -> {
+        { Thick, Blue }
+    }, 
+    Mesh -> {{0}}, 
+    ContourStyle -> Directive[
+            Green, Opacity[0.5], Specularity[White, 30]
+    ], 
+    PlotPoints -> 60,
+    SphericalRegion -> True
+]
+\end{TMcode}
+
+\printindex
 \end{document}

### 2016-05-20 project/article_revised_revised/main.tex
backup

\documentclass{article}

\usepackage[]{amsmath} 
\usepackage[]{amsthm} 
\usepackage[]{amssymb}

\usepackage[]{thmtools} 
\declaretheorem[style=definition, qed=$\diamondsuit$]{definition}
\declaretheorem[style=definition]{corollary}
\declaretheorem[style=definition, qed=$\spadesuit$]{example}
\declaretheorem[style=definition]{proposition}

\usepackage[]{microtype} 
\usepackage[]{enumitem} 
\usepackage[]{graphicx} 
\usepackage[]{subcaption} 
\usepackage[]{makeidx} 
\makeindex

\usepackage[backend=biber, style=alphabetic]{biblatex} 
\addbibresource{mat2000.bib}
\usepackage[]{url} 
\urlstyle{sf}

\usepackage[]{cleveref}

\newcommand{\R}{\ensuremath{\mathbb{R}}}
\newcommand{\C}{\ensuremath{\mathbb{C}}}
\newcommand{\A}{\ensuremath{\mathbb{A}}}
\newcommand{\V}{\ensuremath{\mathbb{V}}}

\renewcommand{\P}{\ensuremath{\mathbb{P}}}

\DeclareMathOperator{\Col}{Col}
\DeclareMathOperator{\rank}{rank}
\DeclareMathOperator{\Bl}{Bl}
\title{Algebraic Surfaces}
\author{Ivar Haugal{\o}kken Stangeby}
\date{\textsc{MAT2000 - Spring 2016}}

\begin{document}
    \tableofcontents

    \section{Preliminaries}
    \label{sec:preliminaries}

    In algebraic geometry the central objects of study is the set of solutions
    to polynomial equations. Finding the zeros of, as an example, the quadratic
    equation
    \begin{equation}
        \notag
        ax^2 + bx + c = 0
    \end{equation}
    is something most people that have been exposed to elementary mathematics
    have done. Later on, one learns how to find the zeroes of more general
    functions, of either one or several variables over some general field of
    coefficients. This leads us to our first definition.

    \begin{definition}[Polynomial ring\index{polynomial ring} \cite{Pol16}]
        A \emph{polynomial ring} in $n$ variables over a field $K$ is the set
        of all functions on the form
        \begin{equation}
            \notag
            f(x_1, \ldots, x_n) = \sum^{}_{\alpha} p_{\alpha}\prod_{i=1}^{n} x_n^{\alpha_i}
        \end{equation}
        where $\alpha = (\alpha_1, \alpha_2, \ldots, \alpha_n)$ is a
        multi-index, and $p_\alpha = p_{\alpha_1,\ldots,\alpha_n}$ is an
        element in $K$.  We denote this polynomial ring $K[x_1, x_2, \ldots,
        x_n]$.
    \end{definition} 
    As an example, the most familiar polynomial ring is $\R[x]$. Elements of
    this ring include
    \begin{align*}
        f_1(x) = x^2, && f_2(x) = \pi, && f_3(x) = x^{17} + x^{3} - e^{4}, && f_4(x) = x^2 + 1.
    \end{align*} 
    Note that some of these have zeroes in the real numbers $\R$ while some
    does not. Most notably is the function $f_4$ whose zeroes lie in the realm
    of complex numbers, $\C$.

    Given a set of polynomials, $f_1, f_2, \ldots, f_m$ in a polynomial ring
    $K[x_1, x_2, \ldots, x_n]$ we can look at their common zeroes. It is in our
    interest to restrict our attention to polynomial rings where the field of
    coefficients is algebraically closed. We recall the definition:

    \begin{definition}[Algebraically closed field {\cite[p.~287]{FK03}}\index{algebraically closed field}]
        A field $K$ is \emph{algebraically closed} if every non-constant
        polynomial in $K[x_1, x_2, \ldots, x_n]$ has a zero in $K$.
    \end{definition}
    Given an algebraically closed field $K$, we define \emph{affine
    $n$-space}\index{affine $n$-space} over $K$ to be the set of all
    $n$-tuples of elements of $K$. We denote affine $n$-space by $\A_K^n$ or
    $\A^n$ if the field of coefficients is clear from context.  For our intents
    and purposes, we will be working over the algebraically closed field of
    complex numbers $\C$.

    We now define the mathematical object that is of utmost importance in the
    study of zeroes of polynomials, namely the variety.
    \begin{definition}[Affine algebraic variety {\cite[p.~3]{Har77}}\index{affine algebraic variety}]
        Given a set of polynomials $f_1, f_2, \ldots, f_m$ from a ring of
        polynomials, the set of points $(x_1, x_2, \ldots, x_n)$ in the affine
        space $\A_K^n$ that satisfy
        \begin{equation}
            \notag
            f_i(x_1, x_2, \ldots, x_n) = 0 \quad \text{for } i = 1, 2, \ldots, m
        \end{equation}
        is called an \emph{affine algebraic variety} and is denoted $\V(f_1,
        f_2, \ldots, f_m)$.
    \end{definition}

    Given the polynomial $f(x) = x^2 - 1$, the affine algebraic variety $\V(f)$
    is then the set $\left\{ 1, -1 \right\}$ since $f(1) = f(-1) = 0$. Given an
    additional polynomial $g(x) = x-1$ we have $\V(f, g) = \left\{ 1 \right\}$
    since $g(1) = 0$, but $g(-1) \neq 0$. Finally, if we also consider the
    polynomial $h(x) = x^2$, then $\V(f, g, h) = \emptyset$. We see as a simple
    fact, and state without proof the following.

    \begin{corollary}
        Given a set of polynomials $f_1, f_2, \ldots, f_m$, the variety
        $\V(f_1, f_2, \ldots, f_m)$ can be expressed as a finite intersection
        \begin{equation}
            \notag
            \V(f_1, \ldots, f_m) = \bigcap_{i=1}^{m} \V(f_i).
        \end{equation}
    \end{corollary}
    It is also of interest, especially when aiming to create and visualize
    interesting figures, that the following hold:
    \begin{corollary}
    Given a set of polynomials $f_1, f_2, \ldots, f_m$, the affine algebraic
    variety of the product $f_1f_2\cdots f_m$ is expressible as a finite union
    \begin{equation}
        \notag
        \V(f_1f_2\cdots f_m) = \bigcup_{i=1}^{m} \V(f_i).
    \end{equation}
    \end{corollary}

    With this in mind, we define one of the more interesting properties of
    varieties, and one we will work with a lot in this article, namely the
    singularities.
    \begin{definition}[Singularity {\cite[p.~3]{Whi08}}\index{singularity}]
        A singularity of the variety $V = \V(f_1, f_2, \ldots, f_m) \subset
        \A_K^n$ is a point $p \in V$ such that the Jacobian matrix
        \begin{equation}
            \notag
            A = \begin{bmatrix}
                \frac{\partial f_1}{\partial x_1} & \hdots & \frac{\partial f_1}{\partial x_n} \\
                \vdots & \ddots & \vdots \\
                \frac{\partial f_m}{\partial x_1} & \hdots & \frac{\partial f_m}{\partial x_n}
            \end{bmatrix}(p)
        \end{equation}
        has rank strictly less than $\min(m, n)$.
    \end{definition}

    A very useful consequence of this definition deal with the varieties of
    single polynomials and will aid us tremendously in the search for
    singularities.
    \begin{corollary}
        \label{crl:single_poly}
        A point $p$ is a singularity of the variety $V = \V(f) \subset \A_K^n$
        if and only if the partial derivatives of $f$ vanish identically at
        $p$. 
    \end{corollary}
    \begin{proof}
        We recall that the rank of a matrix is the dimension of its column
        space, that is for a matrix $A$, $\rank(A) = \dim(\Col(A))$.  Assume
        first that $p$ is a singularity of $V$. Then $\rank(A) < min(n, m) =
        1$. This means that the matrix $A$ has no independent columns,
        otherwise the columns of $A$ would span a space of dimension greater
        than or equal to one. The Jacobian matrix reads in this case
        \begin{equation}
            \notag
            A = \begin{bmatrix}
                \frac{\partial f}{\partial x_1}(p) & \hdots & \frac{\partial f}{\partial x_n}(p)
            \end{bmatrix}.
        \end{equation}
        Since the matrix has no independent columns, we cannot have any pivot
        columns, hence each entry must be zero. Consequently, all partial
        derivatives of $f$ vanish at $p$.  For the opposite implication, assume
        that the partials vanish identically at $p$. Then the Jacobian matrix
        reads
        \begin{equation}
            \notag
            A = \begin{bmatrix}
                0 & \hdots & 0
            \end{bmatrix},
        \end{equation} and consequently $A$ has rank strictly less than one.
        Hence $p$ is a point of singularity of $V$.
    \end{proof}

    We now have the tools needed to start looking at algebraic surfaces.

    \section{Algebraic Surfaces}
    \label{sec:algebraic_surfaces}

    In this section, we try to motivate the definition of an algebraic surface,
    and look at some examples. Hopefully, this will give an intuitive
    introduction to exactly what we will we looking at. We start by looking at
    real algebraic curves, as a good entry point.

    \subsection{Algebraic Curves}
    \label{sub:algebraic_curves}

    From elementary mathematics, one learns about real valued functions,
    $f(x)$, and how to plot these functions by setting $y = f(x)$ and plotting
    points in the $(x, y)$-plane. This yields a curve of sorts, moving around
    in the plane.

    The graph of a function is something of a peculiarity, because it comes
    with some restrictions. Not all curves in the $(x, y)$-plane correspond to
    a specific function. One of the standard methods for verifying whether a
    certain curve corresponds to a function or not, typically taught in school,
    is the \emph{vertical line test}. If you can draw a vertical line through
    two or more points on the curve, then the curve is not associated to any
    specific function. There is however a limit to how interesting
    function-curves can be.

    Having an equation $y = f(x)$ we can form the \emph{equation of a curve at
    zero}. We define a new function in two variables and equate it to zero.
    This yields
    \begin{equation}
        \notag
        g(x, y) = y - f(x) = 0.
    \end{equation}
    The set of points along the curve described by this equation is precisely
    the points that lie in the variety $\V(g(x, y))$.
    If the function $f(x)$ is a polynomial in the variable $x$, i.e., $f(x) \in
    K[x]$, we call the function curve of $f(x)$ algebraic. If the distinction
    between a function graph and curves is to be justified, then there has to
    be curves that are not function graphs.

    Plotting the result when setting $y^2 - x^3$, the plotted curve does not
    pass the vertical line test and is therefore something different from a
    function graph. Similarity, the equation $y^2 = x^3 + x^2$ defines another
    curve that is not a function graph. These are shown in
    \cref{fig:algebraic_curves}, and are prime examples of curves that exhibit
    a singularity. We will get back to the curve shown in \cref{fig:node}
    later.

    \begin{figure}[h!]
        \centering
        \begin{subfigure}[t]{0.45\textwidth}
            \includegraphics[width=\textwidth]{pictures/node.pdf}
            \caption{The curve given by the equation $y^2 - x^3 - x^2 = 0$.
            This curve is commonly known as ''The Node''. It exhibits a
            singularity at the point $p = (0, 0)$ which can be verified by
            computing the partial derivatives at $p$.}
            \label{fig:node}
        \end{subfigure}        
        ~
        \begin{subfigure}[t]{0.45\textwidth}
            \includegraphics[width=\textwidth]{pictures/cusp.pdf}
            \caption{The curve given by $y^2 - x^3 = 0$, or more commonly known
            as ''The Cusp'', is also an example of a singular algebraic curve.
        The point $p = (0, 0)$ is a point of singularity.}
        \end{subfigure}
        \caption{Curves not corresponding to functions.}
        \label{fig:algebraic_curves}
    \end{figure}

    \subsection{Moving up a dimension}
    \label{sub:moving_up_a_dimension}
    With what we have done so far, we can generalize and move up a dimension.
    Instead of considering curves in the $(x, y)$-plane, we can instead look at
    surfaces in the three-dimensional $(x, y, z)$-space. Following the same
    procedure as above, we define a new function $h(x, y, z) = z - g(x, y)$.
    Completely analogous to the definition made for curves, we call the surface
    \emph{algebraic} if $h(x, y, z)$ is a polynomial in the variables $x, y,
    z$. That is $h(x, y, z) \in K[x, y, z]$. We give some examples.

    \begin{example}
        \label{exmpl:algebraic_surfaces}
        The following equations satisfy the definition of an algebraic surface:
        \begin{enumerate}
            \item $h_1(x, y, z) = z - xy = 0$,
            \item $h_2(x, y, z) = z - x^2 - y^2 = 0$,
            \item $h_3(x, y, z) = z - (2x^2 - y)(y - x^2) = 0$,
        \end{enumerate}
        while the following does \emph{not}:
        \begin{enumerate}[resume]
            \item $g_1(x, y, z) = \sin(xy) + z = 0$, and
            \item $g_2(x, y, z) = \tan^2(x) + \pi z y$.
        \end{enumerate}
        These examples are visualized in \cref{fig:algebraic_surfaces}.
        \begin{figure}[]
            \centering
            \begin{subfigure}[t]{0.3\textwidth}
                \includegraphics[height=0.6\textwidth]{pictures/example_one.png} 
                \caption{$h_1(x, y, z) = 0$}
            \end{subfigure}
            ~
            \begin{subfigure}[t]{0.3\textwidth}
                \includegraphics[height=0.6\textwidth]{pictures/example_two.png} 
                \caption{$h_2(x, y, z) = 0$}
            \end{subfigure}
            ~
            \begin{subfigure}[t]{0.3\textwidth}
                \includegraphics[height=0.6\textwidth]{pictures/example_three.png} 
                \caption{$h_3(x, y, z) = 0$}
            \end{subfigure}
            \caption{The surfaces that fell under the category algebraic from
            \cref{exmpl:algebraic_surfaces}.}
            \label{fig:algebraic_surfaces}
        \end{figure}
    \end{example}
    We can of course continue this process of introducing new functions and
    move up a dimension for each time indefinitely.  We complete this section
    by formally defining an algebraic surface. In any case, the surfaces are
    defined by the variety of some defining polynomial.
    \begin{definition}[Algebraic surface\index{algebraic surface}]
        An \emph{algebraic surface} is an algebraic variety of
        dimension\footnote{The dimension of an algebraic variety is a bit hard
            to grasp, and since we only will be dealing with algebraic surfaces
            in this text we do not spend any time classifying the different
        kinds of varieties. The explanation of this is therefore omitted. See
    \cite{Dim16}.}
            two.  We define the \emph{degree}\index{algebraic surface!degree}
            of an algebraic surface to be the highest combined power in the
            defining polynomials terms, and denote this $d$.  That is,
        \begin{equation}
            \notag
            d = \max_{\alpha}\left\{ \alpha_1 + \alpha_2 + \ldots \alpha_n \right\}
        \end{equation}
        where $\alpha_i$ is the power of $x_i$ in a term.
    \end{definition}
    We now take a brief look at some of the properties of these surfaces in
    relation to the singularities.

    \section{Singularities}
    \label{sec:singularities}
    The singularities of an algebraic surface is one of the important elements
    that give the surface its visual appeal. I am sure most people agree with
    me, when I say that a surface full of singularities is more intriguing than
    one with no singularities at all. The singularities, along with their
    configurations, symmetries and their number are things that one can explore
    further.

    \subsection{The number of singularities}
    \label{sub:the_number_of_singularities}

    For an algebraic surface of degree $d$ we let
    $\mu(d)$\index{singularity!$\mu(d)$, maximal number} denote the maximal
    number of singularities such a surface can have.  Determining $\mu(d)$ for
    even relatively small $d$ has proven to be surprisingly difficult.
    \cite{Lab14} gives a very intuitive introduction to the world of
    singularity-hunting and the search for the largest possible $\mu(d)$. We
    shall repeat some of the results here. We start with a non-example of a
    singular surface.

    \begin{proposition}[Singularities of the plane]
        Let $V = \V(f(x, y, z)) = \V(ax + by + cz + d)$ be a plane, then $V$ contains no
        singular points.
    \end{proposition} 
    \begin{proof}
        Assume for contradiction that $p \in V$ is a singular point. Then the
        Jacobian matrix of $ax + by + cz + d$ evaluated at $p$ must have rank
        strictly less than 1, or equivalently, by \cref{crl:single_poly}, the
        partial derivatives vanish at $p$. The partial derivatives of our
        defining polynomial is given by
        \begin{align*}
            \notag
            \frac{\partial f}{\partial x} = a, && \frac{\partial f}{\partial y} = b, &&  \frac{\partial f}{\partial z} = c,
        \end{align*}
        however, none of these are dependent on the point we chose. Hence, for
        these partials to vanish at $p$, we need $a = b = c = 0$. However, our
        variety $V$ then reduces to $\V(d)$ where $d$ is some constant. If $d
        \neq 0$, then $\V(d) = \emptyset$, and therefore does not contain $p$
        at all, which contradicts our assumption. If $d = 0$ on the other hand,
        then we are left with the equation $0 = 0$ which does not describe a
        plane. Consequently, $p$ cannot be a singular point of $V$, and the
        plane therefore exhibits no singularities. 
    \end{proof}
    The argument above tells us that for a general algebraic surface of degree
    $d = 1$, we can have no singularities, hence $\mu(1) = 0$.

    We now look at a surface with four singularities.
    \begin{example}[The Cayley cubic\index{the cayley cubic}]
        The Cayley cubic, a surface of degree $d = 3$, is defined by the
        polynomial
        \begin{equation}
            f(x, y, z) = xyz + (1 - x - y - z)(yz + xz + xy).
        \end{equation}
        We are interested in finding the singularities of this surface, if there are any.
        We first compute the partial derivatives and find that
        \begin{align*}
            \frac{\partial f}{\partial x} &= -(y + z)(2x + y + z - 1), \\
            \frac{\partial f}{\partial y} &= -(x + z)(x + 2y + z - 1), \\
            \frac{\partial f}{\partial z} &= -(x + y)(x + y + 2z - 1).
        \end{align*}
        By inspection, we see that the three partials vanish in the point $(1,
        0, 0)$. Due to the symmetric nature\footnote{The study of symmetric
        functions\index{symmetric function} are also of great interest, which
        we will not go into detail here. However, as a brief mention, given a
        function $f(x_1, x_2, \ldots, x_n)$ and a permutation $\sigma \in S_n$,
        where $S_n$ is the symmetry group on $n$ letters, one can consider the
        set $\mathcal{S} = \left\{ \sigma \in S_n \mid f\left(x_{\sigma(1)},
        x_{\sigma(2)}, \ldots, x_{\sigma(n)}) = f(x_1, x_2, \ldots,
        x_n\right)\right\}$ of permutations that fixes the function $f$. For
        the Cayley cubic discussed above, any $\sigma \in S_3$
        fixes the function, and hence the zeros were easily found by symmetry.} of the
        defining equation, we see that $(0, 1, 0)$ and $(0, 0, 1)$ also are zeroes of
        these polynomials.  Finally, the partials also vanish in the point $(0, 0, 0)$.
        We also see that $f(1, 0, 0) = f(0, 1, 0) = f(0, 0, 1) = f(0, 0, 0) = 0$ so
        these are indeed singularities of the Cayley cubic. The symmetries in the
        defining polynomial is seen reflected in the visualized surface shown in
        \cref{fig:cayley}.

        \begin{figure}[h!]
            \centering
            \includegraphics[width=0.3\linewidth]{pictures/cayley_cubic.png}
            \caption{The Cayley cubic, a degree three surface given by the
                equation $f(x, y, z) = xyz + (1 - x  - y - z)(yz + xz + xy)$.
                It is a good example of a surface whose set of singularities
                show a lot of symmetry. It also exhibit the maximal number of
                singularities for a degree $3$ surface, namely 4.}
            \label{fig:cayley}
        \end{figure}
    \end{example}

    From \cite[p.~7]{Lab14} we have the following relation on the maximum
    number of singularities for $d \geq 3$,
    \begin{equation}
        \notag
        \mu(d) \leq \frac{1}{2}(d(d-1)^2 - 3)
    \end{equation}
    and plugging in $d = 3$ yields $\mu(d) \leq 4$. However, we have already
    presented a surface of degree $d = 3$ with four singularities, so in fact
    $\mu(d) = 4$. The theory of finding the maximal number of singularities of
    a given degree and constructing surfaces that exhibit all is a very
    technical matter. The thesis \cite{oliver2005hypersurfaces} gives a good
    introduction to the construction and visualization of such surfaces.

    We have now had a brief look at singularities of surfaces, and how to find
    them, we now consider how to remove, or \emph{resolve}, such singularities.

    \section{Resolving singularities}
    \label{sec:resolving_singularities}

    We call a variety that exhibits no singularities \emph{smooth}\index{smooth
    variety}. It is often possible to study a smooth copy of a variety with
    singularities if we move from affine space to projective space. This
    process can be defined in many different ways, with the seemingly most
    prevalent one being through maps from various spaces to each other. In this
    paper we will try to keep it as intuitive as possible, by using the more
    geometrically based procedure. We first define the notion of a projective
    space.

    \subsection{Projective space}
    \label{sub:projective_space}

    Informally, projective space generalizes the notion of parallel lines
    intersecting at infinity. The general definition of projective $n$-space is
    the set of lines in $\A^{n+1}$ passing through the origin. For our purposes
    we will take $\A^{n+1} = \C^{n+1}$.

    \begin{definition}[The projective space of dimension $n$\index{projective space}]
        Let $\sim$ be a relation on $\C^{n+1}\setminus \left\{ 0 \right\}$ defined by
        \begin{equation}
            \notag
            (x_1, \ldots, x_{n+1}) \sim (x_1', \ldots, x_{n+1}')
            \iff (x_1', \ldots, x_{n+1}') = \lambda(x_1, \ldots,
            x_{n+1})
        \end{equation}
        where $\lambda$ is some non-zero complex number. This relation can be
        shown to be an equivalence relation. We denote the
        equivalence class of $(x_1, x_2, \ldots, x_{n+1})$ by $\left[ x_1 : x_2
        : \ldots : x_{n+1} \right]$ and define the \emph{projective space of
        dimension $n$} as the set 
        \begin{equation}
            \notag
            \P^{n}_\C = \left\{ \left[ x_1 : \ldots : x_{n+1} \right] \mid (x_1,
            \ldots , x_{n+1}) \in \C^{n+1}\setminus \left\{ 0 \right\}\right\}.
        \end{equation}
        The elements in $\P_\C^n$ are called \emph{homogeneous
        coordinates}\index{homogeneous coordinates}. 
    \end{definition}

    We now turn to the resolution of singularities. There are, as mentioned
    above, many ways of doing this, but we will here apply a geometrical
    approach.

    \subsection{Blowing up a point singularity.}
    \label{sub:blowing_up_a_point_singularity_}

    We start by examining resolutions of single points. Informally, when we
    blow up a point $p$ in $\A^n$, we replace the point $p$ by an entire copy
    of a projective space $\P^{n-1}$ and try to leave the rest of the space
    $\A^n$ ''unchanged''. We will give a few examples, that hopefully will
    illustrate the geometrical aspect of this, but first a formal definition.

    \begin{definition}[Blowup of $\A^{n}$ at a point $p$.]
        We let the \emph{blowup-surface}\index{blowup-surface!in a point}
        $\Bl_p\A^{n}$ be the set of all pairs consisting of a point $p \in
        \A^n$ and a line from $p$ through the origin. The surface is embedded
        in the space $\A^n\times\P^{n-1}$. We write
        \begin{equation}
            \notag
            \Bl_p\A^n = \left\{ (p, L) \mid p \in \A^{n} \text{ and } p \in L \right\}.
        \end{equation}
    \end{definition}
    Now, this is a fairly technical definition, so where do we go from here?
    When blowing up a surface at a point $p$ we are essentially giving the
    surface ''more room'' so that the points creating the singularities can
    pass each other without hassle. In order to aid us with the above
    definition, we show the following claim that tells us under what
    circumstances the point $p$ lie on the line $L$:

    \begin{proposition}
        \label{prop:point_line}
        Given a point $x = (x_1, x_2, \ldots, x_n)$ in $\A^n$ and a point $y =
        [y_1 : y_2 : \ldots : y_n]$ in the space $\P^{n-1}$, the point $x$ is a
        multiple of $y$ if and only if the matrix
        \begin{equation}
            \notag
            A = \begin{bmatrix}
                x_1 & x_2 & \hdots & x_n \\
                y_1 & y_2 & \hdots & y_n
            \end{bmatrix}
        \end{equation}
        has rank less than or equal to one. Similarly, the rank is less than or
        equal to one only if each $2\times 2$-minor $x_iy_j - x_jy_i = 0$ for
        $1 \leq i < j \leq n$.
    \end{proposition} 
    \begin{proof}
        In order to verify this, we recall that $\rank(A) = \rank(A^T)$, so we
        chose to work with the transposed matrix instead. Assume that $x$ is a
        multiple of $y$, so $y = \lambda x$ for some scalar $\lambda$. Then
        \begin{equation}
            \notag
            A^T = \begin{bmatrix}
                x_1 \;\, \lambda x_1 \\
                x_2 \;\, \lambda x_2 \\
                \vdots \\
                x_n \;\, \lambda x_n
            \end{bmatrix}
        \end{equation}
        is the associated matrix. If $x = 0$, then $A^T$ is the zero matrix and
        has therefore rank 0. If $x \neq 0$, then we have exactly one pivot
        column in this matrix, hence $\dim(\Col(A^T)) = 1$. In any case,
        $\rank(A) = \rank(A^T) \leq 1$. For the opposite implication. If $A$
        has rank 0, then $A$ is the zero matrix, and hence trivially, $x$ is a
        multiple of $y$, both being zero. If $A$ has rank $1$, then there is
        one linearly independent column in $A^T$, hence the other column can be
        expressed as a multiple of the first. In other words, $x$ is a multiple
        of $y$.

        If there were a non-zero minor $r\times r$ for $r > 1$, then this
        suggests that there are $r$ linearly independent column vectors, but
        this can't occur since $\rank(A^T) \leq 1$.
    \end{proof}

    With the above result, we can rewrite the definition of the blowup-surface
    $\Bl_p\A^n$ as
    \begin{equation}
        \notag
        \Bl_p\A^n = \left\{ (x_1, x_2, \ldots, x_n; y_1 : y_2 : \ldots : y_n)
    \mid x_iy_j = x_jy_i \text{ for } 1 \leq i < j \leq n \right\}
    \end{equation} which is computationally easier to work with.

    We now give some examples of the computations involved when finding
    blowup-surfaces. We start with the resolution of a singularity in an
    algebraic curve.
    \begin{example}[The node]
        The Node, shown in \cref{fig:node}, is a singular algebraic curve with
        a singularity in the point $p = (0, 0)$. The defining polynomial of the
        node is given by
        \begin{equation}
            \label{eq:node}
            f(x, y) = y^2 - x^2 - x^3,
        \end{equation}
        with the curve $\mathcal{C} = \V(f(x, y, z))$. This curve lies in the
        affine plane $\A^2$, while the resulting blowup-surface of
        $\mathcal{C}$ at the point $(0, 0)$ lies in the space $\A^2 \times
        \P^1$. For a point $(x, y) \in \mathcal{C}$ we need to figure out what
        line passes through both $(x, y)$ and the origin $(0, 0)$. However,
        \cref{prop:point_line} tells us exactly under what circumstances this
        occurs. Let $[u : v]$ be a point in $\P^1$. Then, the matrix
        \begin{equation}
            \notag
            A = \begin{bmatrix}
                x & y \\
                u & v
            \end{bmatrix}
        \end{equation}
        has rank less than or equal to one only when $xv - yu = 0$, our only
        $2\times 2$-minor in this case. This gives us the equation $xv = yu$.
        We chose the point $(u, v)$ as a representative for the equivalence
        class $[u : v]$ however, we might as well chose the point $(u, 1)$ or
        $(1, v)$, since we can set either $u = 1$ or $v = 1$ and then multiply
        the representative by $1 / u$ or $1 / v$ respectively. This makes us
        able to eliminate either $u$ or $v$ from $xv = yu$. This gives us two
        different cases:

        \paragraph{Case 1:}
        \label{par:case_1}

        Setting $u = 1$ we get the equation $y = xv$. This can be substituted
        back into \cref{eq:node} in order to eliminate $y$. This gives us a new
        equation, namely
        \begin{equation}
            \notag
            g(x, v) = x^2(v^2 - 1 - x).
        \end{equation}
        This function $g(x, v)$ is what we call an \emph{affine
        chart}\index{affine chart} of the blowup-surface $\Bl_p\A^2$. It can be
        visualized in the $(x, v)$-plane.

        \paragraph{Case 2:}
        \label{par:case_2}

        Setting $v = 1$ yields the equation $x = yu$ which makes us able to
        eliminate $x$ from \cref{eq:node}. This gives us a new equation
        \begin{equation}
            \notag
            h(y, u) = y^2(1 - u^2 - u^2y).
        \end{equation}
        This is the other affine chart of $\Bl_p\A^2$ and can be visualized in
        the $(y, u)$-plane.  The affine charts given by $h(x, y, z)$ and $g(x,
        y, z)$ are shown visualized in \cref{fig:affine_charts_node}.

        The relation between these two charts, and the original surface can be
        visualized as in \cref{fig:blow_up_intuitive}.

        \begin{figure}[h]
            \centering
            \begin{subfigure}[t]{0.4\textwidth}
                \includegraphics[width=1.1\textwidth]{pictures/node_affine_1.pdf}
                \caption{The affine chart resulting from setting $u = 1$ and
                eliminating $y$ in the original equation.}
            \end{subfigure}
            ~
            \begin{subfigure}[t]{0.4\textwidth}
                \includegraphics[width=1.1\textwidth]{pictures/node_affine_2.pdf}
                \caption{The affine chart resulting from setting $v = 1$ and
                eliminating $x$ in the original equation}
            \end{subfigure}
            \begin{subfigure}[t]{0.4\textwidth}
                \includegraphics[width=1.0\textwidth]{pictures/blow_up_intuitive.pdf}
                \caption{A visual depiction of the blowup surface that clearly
                    shows how the singularity in the origin is resolved. Taken
                from \cite{Whi08}.}
                \label{fig:blow_up_intuitive}
            \end{subfigure}
            \caption{The affine charts of the Node. The vertical line plotted
                in slightly brighter blue, is called the \emph{exceptional
                divisor}\index{exceptional divisor} and corresponds to the
                blowup of what was the origin in the original surface. Note
                that the curves intersect the exceptional divisor in exactly
                two points. This corresponds to the double point singularity in the
                original surface.}
            \label{fig:affine_charts_node}
        \end{figure}
    \end{example}

    Blowing up singular surfaces or curves can be great fun and a lot of
    interesting pictures arise. We now take a look at an example where blowing
    up the surface in a single point is not sufficient for removing the
    singularities.

    \subsection{Blowing up a line of singularity.}
    \label{sub:blowing_up_a_line_of_singularity_}

    In this section we will consider the algebraic surface

    \clearpage
    \printbibliography \printindex
\end{document}

### 2016-05-20 project/visualization/curves/curves.py

@@ -2,22 +2,27 @@ import matplotlib.pyplot
 from numpy import arange
 from numpy import meshgrid
 
+from sys import argv
 
-matplotlib.pyplot.xkcd()
 
-delta = 0.025
-xrange = arange(-5, 5.0, delta)
-yrange = arange(-2.5, 2.5, delta)
+delta = 0.010
+xrange = arange(-2, 2, delta)
+yrange = arange(-2, 2, delta)
 X, Y = meshgrid(xrange,yrange)
 
 # F is one side of the equation, G is the other
 # F = X**2 * (Y**2 - 1 - X)
-F = X**2*(1 - Y**2 - X*Y**3)
+# F = X**2*(1 - Y**2 - X*Y**3)
+# Example 1
+# F = X*Y - X**6 - Y**6
+F = X**2*(1 - Y**2 - X*Y**2)
 G = 0
 
 matplotlib.pyplot.axis('on')
-matplotlib.pyplot.grid('on')
-matplotlib.pyplot.contour(X, Y, (F - G), [0])
+matplotlib.pyplot.grid('off')
+matplotlib.pyplot.contour(X, Y, (F-G), [0])
+matplotlib.pyplot.axvline(x=0)
+matplotlib.pyplot.xlabel('$y$')
+matplotlib.pyplot.ylabel('$u$')
 
-matplotlib.pyplot.savefig('case_1.pdf')
-
+matplotlib.pyplot.savefig('node_affine_2.pdf')

### 2016-05-20 project/visualization/src/implicit_plotter.py

@@ -20,4 +20,4 @@ def implicit_plotter(expr, ext_grid, grid_res=(101, 101, 101),**kwargs):
                        zl:zr:eval('{}j'.format(Nz))]
 
 if __name__ == "__main__":
-    implicit_plotter(expr='x + y + z', ext_grid=(0, 1, 0, 1, 0, 1))
+    implicit_plotter(expr='x - y**2', ext_grid=(0, 1, 0, 1, 0, 1))

### 2016-05-21 project/article_revised_revised/main.tex
work work

@@
 \begin{document}
+    \maketitle
+
+    \begin{abstract}
+       In this paper we give a brief overview of some of the things one can be
+       looking at when studying algebraic surfaces.  Mainly, we look at
+       singularities of such surfaces, how they arise and how one can remove
+       them. The images are rendered using the software \textsc{Surfer} unless
+       specified otherwise.  Some auxilliary \textsc{Python} and
+       \textsc{Mathematica} scripts has also been used.
+    \end{abstract}
+
     \tableofcontents
@@
 
-    In this section we will consider the algebraic surface
+    In this section we will consider the algebraic surface given by the
+    polynomial 
+    \begin{equation}
+        \label{eq:whitney}
+        f(x, y, z) = x^2y - z^2.
+    \end{equation}
+    The associated variety is $\V(f(x, y, z))$. This surface is commonly known
+    as the ''Whitney Umbrella''\index{Whitney umbrella} and is shown visualized in
+    \cref{fig:whitney_umbrella}. As we will see shortly, when trying to blow up
+    this surface in a single point, say the origin, the resulting surface does
+    not seem to be any ''less singular''.
 
+    \begin{figure}[h]
+        \centering
+        \includegraphics[width=0.4\linewidth]{pictures/blowup_line.png}
+        \caption{The ''Whitney Umbrella'' defined by the polynomial $f(x, y, z)
+        = x^2y - z^2$. This surface exhibits singularities all along the
+        $y$-axis, as can easily be pointed out on the picture.}
+        \label{fig:whitney_umbrella}
+    \end{figure}
+
+    \begin{example}[Blowup of the Whitney Umbrella in a single point $p$]
+        If we were to proceed by trying to blow up the origin, we quickly
+        encounter a problem. Following the standard procedure that we used in
+        the previous example, we get the defining equations $xs = yr$, $xt =
+        zr$ and $yt = zs$ with $r, s$ and $t$ our homogeneous coordinates.
+        Contrary to the previous example, we here get three affine charts, one
+        for each homogeneous coordinate.
+
+        \paragraph{Case 1:}
+        Setting $r = 1$ gives us $y = xs$ and $z = xt$. When plugged into
+        \cref{eq:whitney} gives the equation
+        \begin{equation}
+            \notag
+            g(x, s, t) = x^2(xs - t^2) = 0.
+        \end{equation}
+
+        \paragraph{Case 2:}
+        We start by setting $s = 1$, which gives us $x = yr$ and $z = yt$.
+        Eliminating $x$ and $y$ from \cref{eq:whitney} yields the equation
+        \begin{equation}
+            \notag
+            h(y, r, t) = y^2(yr^2 - t^2) = 0.
+        \end{equation}
+
+        \paragraph{Case 3:}
+        Finally, setting $t = 1$ yields $x = zr$ and $y = zs$, giving an
+        equation
+        \begin{equation}
+            \notag
+            i(z, r, s) = z^2(r^2sz - 1) = 0.
+        \end{equation}
+        These can be visualized, and the surfaces are shown in
+        \cref{fig:blowup_line_charts}. As we see in \cref{fig:whitney_blowup} 
+        the blowup fails to remove the singularities.
+
+        \begin{figure}[h]
+            \centering
+            \begin{subfigure}[t]{0.3\textwidth}
+                \includegraphics[width=\textwidth]{pictures/blowup_line_2.png}
+                \caption{$g(x, s, t) = 0$}
+            \end{subfigure}      
+            ~
+            \begin{subfigure}[t]{0.3\textwidth}
+                \includegraphics[width=\textwidth]{pictures/blowup_line_1.png}
+                \caption{$h(y, r, t) = 0$}
+                \label{fig:whitney_blowup}
+            \end{subfigure}      
+            ~
+            \begin{subfigure}[t]{0.3\textwidth}
+                \includegraphics[width=\textwidth]{pictures/blowup_line_3.png}
+                \caption{$i(z, r, s) = 0$}
+            \end{subfigure}      
+        \caption{The affine charts of the blowup of the Whitney Umbrella at the
+            point $p = (0, 0, 0)$. As we can see the singularities improve in
+            two of the charts, and in one case the chart is smooth. However,
+            the middle surface is identical to our original surface, so the
+            blow-up has failed to resolve all the singularities of the original
+            variety.}
+            \label{fig:blowup_line_charts}
+        \end{figure}      
+    \end{example}
+
+    As seen, in this case it is not sufficient to blow up only in one single
+    point. We may have to blow up the surface at a singular line, singular
+    curve or in general, any other set. We make a definition of the blowup of a
+    space in a line $L$, analogous to the blowup in a point,  We specialize
+    this definition for $\A^3$.
+
+    \begin{definition}[Blowup of $\A^3$ at a line $L$]
+        We let the blowup-surface $\Bl_L\A^3$ be the set of all pairs
+        consisting of a point $p \in \A^3$ and a point $\lambda \in \P^1$ such
+        that $H_\lambda$ is the unique plane that contains both the point $p$
+        and the line $L$ and write
+        \begin{equation}
+            \notag
+            \Bl_L\A^3 = \left\{ (p, \lambda) \in \A^3 \times \P^1 \mid p, L \in H_\lambda \right\}.
+        \end{equation}
+    \end{definition}
+
+    We illustrate this definition by again considering the Whitney Umbrella.
+    \begin{example}[Blowup of the Whitney Umbrella at a line $L$]
+        The singularities of the Whitney Umbrella lie along the $y$-axis, which
+        is described by the equations $x = z = 0$. Now, our goal here is to
+        find under what circumstances a plane $H_\lambda$ contains the point $p
+        = (x, y, z)$ as well as the line $L$. In other words, we want both the
+        $y$-axis and $p$ to lie in the plane $rx + sy + tz = 0$.
+
+        We first find the planes that contain the $y$-axis. Choosing the vector
+        $(0, 1, 0)$ as a representative for the $y$-axis we compute the cross
+        product between this vector and the point $(r, s, t)$ as a
+        representative for the plane.  This yields a normal vector given by
+        \begin{equation}
+            \notag
+            \vec{n} = \begin{vmatrix}
+                \hat{i} & \hat{j} & \hat{k} \\
+                r & s & t \\
+                0 & 1 & 0
+            \end{vmatrix} = (-t, 0, r).
+        \end{equation}
+        A plane containing the $y$-axis is then given by the equation $-tx + 0y
+        + rz = rz - tx = 0$. We can now consider the two affine charts by
+        setting $t = 1$ and $s = 1$ respectively.
+
+        \paragraph{Case 1:} 
+        \label{par:case_1_}
+        Setting $r = 1$ yields $z = tx$ which when plugged into $x^2y - z^2 = 0$
+        gives us $x^2y - (t^2x^2) = x^2(y - t^2) = 0$. This affine chart can be
+        visualized in the $(x, y, t)$-plane.
+
+        \paragraph{Case 2:}
+        \label{par:case_2_}
+        Setting $t = 1$ yields $x = rz$ which when plugged into $x^2y-z^2 = 0$
+        gives the equation $r^2z^2y - z^2 = r^2(z^2y - 1) = 0$. This can be
+        visualized in the $(r, y, z)$-plane.
+
+        \begin{figure}[h!]
+            \centering
+            \begin{subfigure}[t]{0.3\textwidth}
+                \includegraphics[width=\textwidth]{pictures/line_blowup_affine_1.png}
+                \caption{The surface given by the equation $x^2(y - t^2) = 0$. The
+                    curve of intersection between this surface and the exceptional
+                    divisor looks like it can be ''pinched'' into a double line, i.e., the
+                original line of singularity.}
+            \end{subfigure}
+            ~
+            \begin{subfigure}[t]{0.3\textwidth}
+                \includegraphics[width=\textwidth]{pictures/line_blowup_affine_2.png} 
+                \caption{The surface given by the equation $r^2(z^2y - 1) = 0$. The
+                    intersection between this surface and the exceptional divisor can
+                    also be ''fused'' together to form a double line, by laying the
+                surfaces flat on top of each other.}
+            \end{subfigure}
+            \\
+            \begin{subfigure}[t]{0.3\textwidth}
+                \includegraphics[width=\textwidth]{pictures/affine_1_exceptional.pdf} 
+            \end{subfigure}
+            ~
+            \begin{subfigure}[t]{0.3\textwidth}
+                \includegraphics[width=\textwidth]{pictures/affine_2_exceptional.pdf} 
+            \end{subfigure}
+            \caption{The two affine charts of the variety $\V(x^2y-z^2)$. The top
+                row is rendered by \textsc{Surfer}, and the bottom row rendered by
+                \textsc{Mathematica.} The exceptional divisor in the top row has
+                been divided out and is therefore not shown because \textsc{Surfer}
+                does a poor job at rendering planes. As seen, the charts are smooth and
+                exhibits no singularities. The intersection with the exceptional divisor is
+            rendered in blue in the bottom row.}
+        \end{figure}
+    \end{example}
+
+    While in the above example we blew up a very specific line of
+    singularities, namely the $y$-axis, it can be of interest to talk a little
+    bit about how to blow up general lines. It all boils down to generalizing
+    the construction of the plane for a more general line $L$. Assuming $L$ is
+    parametrized by $x = at, y = bt, z = ct$, with $a, b$ and $c$ non-zero, we
+    can form the symmetric equations 
+    \begin{equation}
+        \notag
+        t = \frac{x}{a} = \frac{y}{b} = \frac{z}{c}.
+    \end{equation}
+    The line $L$ is then completely determined by two of the three possible
+    relations $xb - ya = 0$, $xc - za = 0$ and $yc - zb = 0$.  In order to find
+    the respective plane, we are interested in knowing when the plane $rx + sy
+    + tz = 0$ can be expressed in terms of the line given by $xb - ya = 0$ and
+    $yc - zb = 0$. This can be solved as a matrix problem. Given the matrix
+    \begin{equation}
+        \notag
+        A = \begin{bmatrix}
+            b & 0 & r \\
+            -a & c & s \\
+            0 & -b & t
+        \end{bmatrix}
+    \end{equation}
+    one of the columns of $A$ is linearly dependent of the two others if
+    $\det(A) = 0$. Solving $\det(A) = 0$ gives the equation
+    \begin{equation}
+        \notag
+        \det(A) = b(ct + sb) + rab = 0.
+    \end{equation}
+    We can now solve for either one of $r, s$ and $t$. Solving for $r$ we get
+    $r = -( ct + sb ) / a$ which we can substitute back into the equation for
+    our plane.  The plane is then determined by 
+    \begin{equation}
+        \notag
+        \frac{-ct - sb}{a}x + sy + tz = 0.
+    \end{equation}
+    Multiplying through by $a$ we achieve
+    \begin{align*}
+        (-sb - ct)x + sya + tza &= -sbx - ctx + sya + tza \\
+                                &= t(za - cx) + s(ya - bx) = 0,
+    \end{align*}
+    which gives us the restrictions that we need.
     \clearpage

### 2016-05-22 project/article_revised_revised/main.tex
fixed some of the comments

@@
 
+\input{code_style.tex}
 \usepackage[]{microtype} 
@@
 
-\usepackage[]{cleveref}
+\usepackage[noabbrev]{cleveref}
 
@@
     functions, of either one or several variables over some general field of
-    coefficients. This leads us to our first definition.
+    coefficients. Of special interest to us, are the polynomials over one or
+    several variables, given as elements of a \emph{polynomial ring} over some
+    field which we denote $K[x_1, x_2, \ldots, x_n]$.
 
-    \begin{definition}[Polynomial ring\index{polynomial ring} \cite{Pol16}]
-        A \emph{polynomial ring} in $n$ variables over a field $K$ is the set
-        of all functions on the form
-        \begin{equation}
-            \notag
-            f(x_1, \ldots, x_n) = \sum^{}_{\alpha} p_{\alpha}\prod_{i=1}^{n} x_n^{\alpha_i}
-        \end{equation}
-        where $\alpha = (\alpha_1, \alpha_2, \ldots, \alpha_n)$ is a
-        multi-index, and $p_\alpha = p_{\alpha_1,\ldots,\alpha_n}$ is an
-        element in $K$.  We denote this polynomial ring $K[x_1, x_2, \ldots,
-        x_n]$.
-    \end{definition} 
     As an example, the most familiar polynomial ring is $\R[x]$. Elements of
@@
     interest to restrict our attention to polynomial rings where the field of
-    coefficients is algebraically closed. We recall the definition:
+    coefficients is algebraically closed.
 
-    \begin{definition}[Algebraically closed field {\cite[p.~287]{FK03}}\index{algebraically closed field}]
-        A field $K$ is \emph{algebraically closed} if every non-constant
-        polynomial in $K[x_1, x_2, \ldots, x_n]$ has a zero in $K$.
-    \end{definition}
     Given an algebraically closed field $K$, we define \emph{affine
-    $n$-space}\index{affine $n$-space} over $K$ to be the set of all
-    $n$-tuples of elements of $K$. We denote affine $n$-space by $\A_K^n$ or
-    $\A^n$ if the field of coefficients is clear from context.  For our intents
-    and purposes, we will be working over the algebraically closed field of
-    complex numbers $\C$.
+    $n$-space}\index{affine $n$-space} over $K$ to be the set of all $n$-tuples
+    of elements of $K$. We denote affine $n$-space by $\A_K^n$ or $\A^n$ if the
+    field of coefficients is clear from context.  For our intents and purposes,
+    we will be working over the algebraically closed field of complex numbers
+    $\C$.
 
@@
     varieties, and one we will work with a lot in this article, namely the
-    singularities.
+    singularities. A singular point is, strictly speaking, a point on a surface
+    where the behaviour of the surface is slightly strange, say the surface
+    passes through itself. We are often interested in studying how a surface
+    behaves locally near a point $p$. We do this using \emph{linearization at
+    $p$}. This is what we do when we consider the tangent line to a point for a
+    single-valued function, the tangent plane for a general surface in three
+    dimensions, and a tangent \emph{hyperplane} for general surfaces in $n$
+    dimensions. The linearization of a surface given by $f(x_1, x_2, \ldots,
+    x_n) = 0$, at a point $p_0$ is given by
+    \begin{equation}
+        \notag
+        L(p) = f(p_0) + J_f(p_0)(p - p_0)
+    \end{equation}
+    where $J_f(p_0)$ is the Jacobian matrix of $f(x_1, x_2, \ldots, x_n)$
+    evaluated at $p_0$. If this matrix has rank less than $n$, then the
+    linearization is not uniquely defined, i.e., there are more than one plane
+    satisfying the linearization. This leads us to the next defintion.
+
     \begin{definition}[Singularity {\cite[p.~3]{Whi08}}\index{singularity}]
@@
         first that $p$ is a singularity of $V$. Then $\rank(A) < min(n, m) =
-        1$. This means that the matrix $A$ has no independent columns,
-        otherwise the columns of $A$ would span a space of dimension greater
-        than or equal to one. The Jacobian matrix reads in this case
-        \begin{equation}
-            \notag
-            A = \begin{bmatrix}
-                \frac{\partial f}{\partial x_1}(p) & \hdots & \frac{\partial f}{\partial x_n}(p)
-            \end{bmatrix}.
-        \end{equation}
-        Since the matrix has no independent columns, we cannot have any pivot
-        columns, hence each entry must be zero. Consequently, all partial
-        derivatives of $f$ vanish at $p$.  For the opposite implication, assume
-        that the partials vanish identically at $p$. Then the Jacobian matrix
-        reads
+        1$. Consequently, we must have $\rank(A) = 0$, and therefore $A$ equal
+        to the zero matrix. All the partial derivatives are therefore zero.
+        For the opposite implication, assume that the partials vanish
+        identically at $p$. Then the Jacobian matrix reads
         \begin{equation}
@@
             \end{bmatrix},
-        \end{equation} and consequently $A$ has rank strictly less than one.
-        Hence $p$ is a point of singularity of $V$.
+        \end{equation}
+        and consequently $A$ has rank strictly less than one. Hence $p$ is a
+        point of singularity of $V$.
     \end{proof}
+
+    Finally, we want to define the notion of a projective space, as this will
+    aid us in the resolution of singularities later on.  Informally, projective
+    space generalizes the notion of parallel lines intersecting at infinity.
+    The general definition of projective $n$-space is the set of lines in
+    $\A^{n+1}$ passing through the origin. For our purposes we will take
+    $\A^{n+1} = \C^{n+1}$.
+
+    \begin{definition}[The projective space of dimension $n$\index{projective space}]
+        Let $\sim$ be a relation on $\C^{n+1}\setminus \left\{ 0 \right\}$ defined by
+        \begin{equation}
+            \notag
+            (x_1, \ldots, x_{n+1}) \sim (x_1', \ldots, x_{n+1}')
+            \iff (x_1', \ldots, x_{n+1}') = \lambda(x_1, \ldots,
+            x_{n+1})
+        \end{equation}
+        where $\lambda$ is some non-zero complex number. This relation can be
+        shown to be an equivalence relation. We denote the
+        equivalence class of $(x_1, x_2, \ldots, x_{n+1})$ by $\left[ x_1 : x_2
+        : \ldots : x_{n+1} \right]$ and define the \emph{projective space of
+        dimension $n$} as the set 
+        \begin{equation}
+            \notag
+            \P^{n}_\C = \left\{ \left[ x_1 : \ldots : x_{n+1} \right] \mid (x_1,
+            \ldots , x_{n+1}) \in \C^{n+1}\setminus \left\{ 0 \right\}\right\}.
+        \end{equation}
+        The elements in $\P_\C^n$ are called \emph{homogeneous
+        coordinates}\index{homogeneous coordinates}. 
+    \end{definition}
 
@@
 
-    \section{Algebraic Surfaces}
-    \label{sec:algebraic_surfaces}
+    \section{Introduction to algebraic surfaces}
 
@@
             \caption{The curve given by the equation $y^2 - x^3 - x^2 = 0$.
-            This curve is commonly known as ''The Node''. It exhibits a
+            This curve is commonly known as ``The Node''. It exhibits a
             singularity at the point $p = (0, 0)$ which can be verified by
@@
             \caption{The curve given by $y^2 - x^3 = 0$, or more commonly known
-            as ''The Cusp'', is also an example of a singular algebraic curve.
+            as ``The Cusp'', is also an example of a singular algebraic curve.
         The point $p = (0, 0)$ is a point of singularity.}
@@
 
-    From \cite[p.~7]{Lab14} we have the following relation on the maximum
+    From \cite[p.~7]{Lab14} we have the following upper bound on the maximum
     number of singularities for $d \geq 3$,
@@
     paper we will try to keep it as intuitive as possible, by using the more
-    geometrically based procedure. We first define the notion of a projective
-    space.
-
-    \subsection{Projective space}
-    \label{sub:projective_space}
-
-    Informally, projective space generalizes the notion of parallel lines
-    intersecting at infinity. The general definition of projective $n$-space is
-    the set of lines in $\A^{n+1}$ passing through the origin. For our purposes
-    we will take $\A^{n+1} = \C^{n+1}$.
-
-    \begin{definition}[The projective space of dimension $n$\index{projective space}]
-        Let $\sim$ be a relation on $\C^{n+1}\setminus \left\{ 0 \right\}$ defined by
-        \begin{equation}
-            \notag
-            (x_1, \ldots, x_{n+1}) \sim (x_1', \ldots, x_{n+1}')
-            \iff (x_1', \ldots, x_{n+1}') = \lambda(x_1, \ldots,
-            x_{n+1})
-        \end{equation}
-        where $\lambda$ is some non-zero complex number. This relation can be
-        shown to be an equivalence relation. We denote the
-        equivalence class of $(x_1, x_2, \ldots, x_{n+1})$ by $\left[ x_1 : x_2
-        : \ldots : x_{n+1} \right]$ and define the \emph{projective space of
-        dimension $n$} as the set 
-        \begin{equation}
-            \notag
-            \P^{n}_\C = \left\{ \left[ x_1 : \ldots : x_{n+1} \right] \mid (x_1,
-            \ldots , x_{n+1}) \in \C^{n+1}\setminus \left\{ 0 \right\}\right\}.
-        \end{equation}
-        The elements in $\P_\C^n$ are called \emph{homogeneous
-        coordinates}\index{homogeneous coordinates}. 
-    \end{definition}
+    geometrically based procedure.
 
@@
     of a projective space $\P^{n-1}$ and try to leave the rest of the space
-    $\A^n$ ''unchanged''. We will give a few examples, that hopefully will
+    $\A^n$ ``unchanged''. We will give a few examples, that hopefully will
     illustrate the geometrical aspect of this, but first a formal definition.
@@
     When blowing up a surface at a point $p$ we are essentially giving the
-    surface ''more room'' so that the points creating the singularities can
+    surface ``more room'' so that the points creating the singularities can
     pass each other without hassle. In order to aid us with the above
@@
         \end{equation}
-        has rank less than or equal to one. Similarly, the rank is less than or
-        equal to one only if each $2\times 2$-minor $x_iy_j - x_jy_i = 0$ for
+        has rank less than or equal to 1. Similarly, the rank is less than or
+        equal to 1 if and only if each $2\times 2$-minor $x_iy_j - x_jy_i = 0$ for
         $1 \leq i < j \leq n$.
@@
             \notag
-            g(x, v) = x^2(v^2 - 1 - x).
+            g(x, v) = x^2(v^2 - 1 - x) = 0.
         \end{equation}
-        This function $g(x, v)$ is what we call an \emph{affine
+        The variety $\V(g(x, v))$ is what we call an \emph{affine
         chart}\index{affine chart} of the blowup-surface $\Bl_p\A^2$. It can be
@@
             \notag
-            h(y, u) = y^2(1 - u^2 - u^2y).
+            h(y, u) = y^2(1 - u^2 - u^2y) = 0.
         \end{equation}
@@
 
-        \begin{figure}[h]
+        \begin{figure}[h!]
             \centering
@@
     The associated variety is $\V(f(x, y, z))$. This surface is commonly known
-    as the ''Whitney Umbrella''\index{Whitney umbrella} and is shown visualized in
+    as the ``Whitney Umbrella''\index{Whitney umbrella} and is shown visualized in
     \cref{fig:whitney_umbrella}. As we will see shortly, when trying to blow up
     this surface in a single point, say the origin, the resulting surface does
-    not seem to be any ''less singular''.
+    not seem to be any ``less singular''.
 
@@
         \includegraphics[width=0.4\linewidth]{pictures/blowup_line.png}
-        \caption{The ''Whitney Umbrella'' defined by the polynomial $f(x, y, z)
+        \caption{The ``Whitney Umbrella'' defined by the polynomial $f(x, y, z)
         = x^2y - z^2$. This surface exhibits singularities all along the
@@
 
+        The two affine charts are shown visualized in \cref{fig:affine_line}.
+
         \begin{figure}[h!]
@@
                     curve of intersection between this surface and the exceptional
-                    divisor looks like it can be ''pinched'' into a double line, i.e., the
+                    divisor looks like it can be ``pinched'' into a double line, i.e., the
                 original line of singularity.}
@@
                     intersection between this surface and the exceptional divisor can
-                    also be ''fused'' together to form a double line, by laying the
+                    also be ``fused'' together to form a double line, by laying the
                 surfaces flat on top of each other.}
@@
             rendered in blue in the bottom row.}
+            \label{fig:affine_line}
         \end{figure}
@@
     which gives us the restrictions that we need.
+
     \clearpage 
-    \printbibliography \printindex
+    \appendix
+
+    \section{Code snippets}
+    \label{sec:code_snippets}
+
+    \lstset{deletekeywords={xrange}}
+    \begin{TMcode}{python}{curves.py - \url{http://stackoverflow.com/a/2484594}}{Plotting algebraic curves}
+    import matplotlib.pyplot
+    from numpy import arange
+    from numpy import meshgrid
+
+    delta = 0.025
+    xrange = arange(-5, 5.0, delta)
+    yrange = arange(-2.5, 2.5, delta)
+    X, Y = meshgrid(xrange,yrange)
+
+    # F is one side of the equation, G is the other
+    F = X**2*(1 - Y**2 - X*Y**3)
+    G = 0
+
+    matplotlib.pyplot.axis('on')
+    matplotlib.pyplot.grid('on')
+    matplotlib.pyplot.contour(X, Y, (F - G), [0])
+
+    matplotlib.pyplot.savefig('output.pdf')
+    \end{TMcode}
+
+    \clearpage
+    \begin{TMcode}{mathematica}{intersection.nb - \url{http://community.wolfram.com/groups/-/m/t/178218?p_p_auth=lT7er3Rd}}{Plotting the intersection between to surfaces}
+
+    (* affine chart and exceptional divisor *)
+    h = z^2 y - 1;
+    g = x^2;
+
+    ContourPlot3D[
+        {h == 0, g == 0}, 
+        (* bounding box dimensions *)
+        {x, -3, 3}, 
+        {y, -3, 3}, 
+        {z, -3, 3}, 
+        MeshFunctions -> {
+            Function[ {x, y, z, f}, h - g ]
+        }, 
+        MeshStyle -> {
+            { Thick, Blue }
+        }, 
+        Mesh -> {{0}}, 
+        ContourStyle -> Directive[
+                Green, Opacity[0.5], Specularity[White, 30]
+        ], 
+        PlotPoints -> 60,
+        SphericalRegion -> True
+    ]
+    \end{TMcode}
+
+    \printbibliography 
 \end{document}

### 2016-05-24 project/article_revised_revised/main.tex
final

@@
 
+\usepackage[]{fontspec} 
 \input{code_style.tex}
@@
 \usepackage[]{makeidx}
+
+\usepackage[margin=2cm]{geometry} 
 \makeindex
@@
 
-\usepackage[noabbrev]{cleveref}
-
 \newcommand{\R}{\ensuremath{\mathbb{R}}}
@@
 \DeclareMathOperator{\Bl}{Bl}
+
+\usepackage[]{hyperref} 
+\usepackage[noabbrev]{cleveref} 
 \title{Algebraic Surfaces}
@@
        In this paper we give a brief overview of some of the things one can be
-       looking at when studying algebraic surfaces.  Mainly, we look at
-       singularities of such surfaces, how they arise and how one can remove
-       them. The images are rendered using the software \textsc{Surfer} unless
-       specified otherwise. Some auxilliary \textsc{Python} and
-       \textsc{Mathematica} scripts has also been used.
+       looking at when studying algebraic surfaces. We use a fairly naive
+       approach without the very abstract mathematical machinery that this
+       field has to offer. In \cref{sec:preliminaries} we give some of the
+       needed mathematical preliminaries. \Cref{sec:intro} motivates the
+       definition of an algebraic surface, the main object of study. In
+       \cref{sec:singularities} we look at one of the main properties of
+       algebraic surfaces, namely the singular points, and proceed by showing
+       how to resolve them i \cref{sec:resolving_singularities}. Finally, we
+       give an example of how to construct a singular surface with a certain
+       symmetry in \cref{sec:constructing_singular_surfaces}.  The images are
+       rendered using the software \textsc{Surfer} unless specified otherwise.
+       Some auxiliary \textsc{Python} and \textsc{Mathematica} scripts has also
+       been used. 
     \end{abstract}
@@
     \begin{corollary}
+        \label{crl:set_one}
         Given a set of polynomials $f_1, f_2, \ldots, f_m$, the variety
@@
     \begin{corollary}
+        \label{crl:set_two}
         Given a set of polynomials $f_1, f_2, \ldots, f_m$, the affine algebraic
@@
     linearization is not uniquely defined, i.e., there are more than one plane
-    satisfying the linearization. This leads us to the next defintion.
+    satisfying the linearization. This leads us to the next definition.
 
@@
     \begin{definition}[The projective space of dimension $n$\index{projective space}]
-        Let $\sim$ be a relation on $\C^{n+1}\setminus \left\{ 0 \right\}$ defined by
+        Let $\sim$ be a relation on $\C^{n+1} \setminus \left\{ 0 \right\}$ defined by
         \begin{equation}
@@
     \section{Introduction to algebraic surfaces}
-
+    \label{sec:intro}    
     In this section, we try to motivate the definition of an algebraic surface,
@@
     points in the $(x, y)$-plane. This yields a curve of sorts, moving around
-    in the plane.
-
-    The graph of a function is something of a peculiarity, because it comes
-    with some restrictions. Not all curves in the $(x, y)$-plane correspond to
-    a specific function. One of the standard methods for verifying whether a
-    certain curve corresponds to a function or not, typically taught in school,
-    is the \emph{vertical line test}. If you can draw a vertical line through
-    two or more points on the curve, then the curve is not associated to any
-    specific function. There is however a limit to how interesting
-    function-curves can be.
-
-    Having an equation $y = f(x)$ we can form the \emph{equation of a curve at
-    zero}. We define a new function in two variables and equate it to zero.
-    This yields
+    in the plane.  Having an equation $y = f(x)$ we can form the \emph{equation
+    of a curve at zero}. We define a new function in two variables and equate
+    it to zero. This yields
     \begin{equation}
@@
     The set of points along the curve described by this equation is precisely
-    the points that lie in the variety $\V(g(x, y))$.
-    If the function $f(x)$ is a polynomial in the variable $x$, i.e., $f(x) \in
-    K[x]$, we call the function curve of $f(x)$ algebraic. If the distinction
-    between a function graph and curves is to be justified, then there has to
-    be curves that are not function graphs.
+    the points that lie in the variety $\V(g(x, y))$.  If the function $f(x)$
+    is a polynomial in the variable $x$, i.e., $f(x) \in K[x]$, we call the
+    function curve of $f(x)$ algebraic. Now, all the curves formed in this way
+    corresponds to some function. There are however curves that does not
+    correspond to functions and these are given as general polynomials.
 
@@
         \end{subfigure}
-        \caption{Curves not corresponding to functions.}
+        \caption{First examples of algebraic curves. Visualized using
+        \textsc{Python}. Code can be seen in \cref{sec:code_snippets}.}
         \label{fig:algebraic_curves}
@@
     procedure as above, we define a new function $h(x, y, z) = z - g(x, y)$.
-    Completely analogous to the definition made for curves, we call the surface
-    \emph{algebraic} if $h(x, y, z)$ is a polynomial in the variables $x, y,
-    z$. That is $h(x, y, z) \in K[x, y, z]$. We give some examples.
+    This function equated at zero yields a surface.  Completely analogous to
+    the definition made for curves, we call the surface \emph{algebraic} if
+    $h(x, y, z)$ is a polynomial in the variables $x, y, z$. That is $h(x, y,
+    z) \in K[x, y, z]$. We give some examples.
 
@@
     \begin{definition}[Algebraic surface\index{algebraic surface}]
-        An \emph{algebraic surface} is an algebraic variety of
-        dimension\footnote{The dimension of an algebraic variety is a bit hard
-            to grasp, and since we only will be dealing with algebraic surfaces
-            in this text we do not spend any time classifying the different
-        kinds of varieties. The explanation of this is therefore omitted. See
-    \cite{Dim16}.}
-            two.  We define the \emph{degree}\index{algebraic surface!degree}
-            of an algebraic surface to be the highest combined power in the
-            defining polynomials terms, and denote this $d$.  That is,
+        Given a set of polynomials $f_1, f_2, \ldots, f_m$ in three variables,
+        i.e., in $K[x, y, z]$ we define an \emph{algebraic surface} to be the
+        algebraic variety $\V(f_1, f_2, \ldots, f_m)$.  We define the
+        \emph{degree}\index{algebraic surface!degree} of an algebraic surface
+        to be the highest combined power in the defining polynomials terms, and
+        denote this $d$. That is,
         \begin{equation}
@@
     \label{sec:singularities}
-    The singularities of an algebraic surface is one of the important elements
-    that give the surface its visual appeal. I am sure most people agree with
-    me, when I say that a surface full of singularities is more intriguing than
-    one with no singularities at all. The singularities, along with their
-    configurations, symmetries and their number are things that one can explore
-    further.
+
+    When studying algebraic surfaces we typically differentiate between two
+    types of surfaces, namely the singular ones and the non-singular ones.
+    These two categories are each interesting in their own respect. A lot of
+    mathematical theory concerning surfaces only hold for surfaces that exhibit
+    no singularities at all. It is therefore often of interest to somehow
+    remove these singularities from the surface in order to apply known
+    results.
+
+    The theory of singular surfaces on the other hand contains a lot of
+    interesting theory that deals with: the classification of singularities,
+    what different types of singularities are there; resolution of
+    singularities; and deformation theory.
 
@@
     \end{equation}
-    and plugging in $d = 3$ yields $\mu(d) \leq 4$. However, we have already
+    and plugging in $d = 3$ yields $\mu(3) \leq 4$. However, we have already
     presented a surface of degree $d = 3$ with four singularities, so in fact
@@
 
-    We now turn to the resolution of singularities. There are, as mentioned
-    above, many ways of doing this, but we will here apply a geometrical
-    approach.
-
-    \subsection{Blowing up a point singularity.}
+    When we blow up, we blow up along a \emph{center} of the variety. By making
+    sure that this center is contained in the singular locus of the variety, we
+    do not change the surrounding surface. We first consider the case where the
+    blowup center is a single point, and then an example where the center is a
+    line.
+
+    \subsection{Blowing up a point}
     \label{sub:blowing_up_a_point_singularity_}
@@
     of a projective space $\P^{n-1}$ and try to leave the rest of the space
-    $\A^n$ ``unchanged''. We will give a few examples, that hopefully will
-    illustrate the geometrical aspect of this, but first a formal definition.
+    $\A^n$ ``unchanged''. We will
+    give a few examples, that hopefully will illustrate the geometrical aspect
+    of this, but first a formal definition.
 
@@
         the $(y, u)$-plane.  The affine charts given by $h(x, y, z)$ and $g(x,
-        y, z)$ are shown visualized in \cref{fig:affine_charts_node}.
-
-        The relation between these two charts, and the original surface can be
+        y, z)$ are shown visualized in \cref{fig:affine_charts_node}.  The
+        relation between these two charts, and the original surface can be
         visualized as in \cref{fig:blow_up_intuitive}.
@@
             \centering
-            \begin{subfigure}[t]{0.4\textwidth}
+            \begin{subfigure}[t]{0.3\textwidth}
                 \includegraphics[width=1.1\textwidth]{pictures/node_affine_1.pdf}
@@
             ~
-            \begin{subfigure}[t]{0.4\textwidth}
+            \begin{subfigure}[t]{0.3\textwidth}
                 \includegraphics[width=1.1\textwidth]{pictures/node_affine_2.pdf}
@@
                 eliminating $x$ in the original equation}
-            \end{subfigure}
-            \begin{subfigure}[t]{0.4\textwidth}
+            \end{subfigure} \\
+            \begin{subfigure}[t]{0.3\textwidth}
                 \includegraphics[width=1.0\textwidth]{pictures/blow_up_intuitive.pdf}
@@
                 blowup of what was the origin in the original surface. Note
-                that the curves intersect the exceptional divisor in exactly
-                two points. This corresponds to the double point singularity in the
-                original surface.}
+                that the curves intersect the exceptional divisor in exactly two
+                points. This corresponds to the double point singularity in the
+                original surface. The top row is visualized in \textsc{Python}.}
             \label{fig:affine_charts_node}
@@
 
-    \subsection{Blowing up a line of singularity.}
+    \subsection{Blowing up a line}
     \label{sub:blowing_up_a_line_of_singularity_}
@@
         \centering
-        \includegraphics[width=0.4\linewidth]{pictures/blowup_line.png}
+        \includegraphics[width=0.3\linewidth]{pictures/blowup_line.png}
         \caption{The ``Whitney Umbrella'' defined by the polynomial $f(x, y, z)
@@
             \centering
-            \begin{subfigure}[t]{0.3\textwidth}
+            \begin{subfigure}[t]{0.2\textwidth}
                 \includegraphics[width=\textwidth]{pictures/blowup_line_2.png}
@@
             ~
-            \begin{subfigure}[t]{0.3\textwidth}
+            \begin{subfigure}[t]{0.2\textwidth}
                 \includegraphics[width=\textwidth]{pictures/blowup_line_1.png}
@@
             ~
-            \begin{subfigure}[t]{0.3\textwidth}
+            \begin{subfigure}[t]{0.2\textwidth}
                 \includegraphics[width=\textwidth]{pictures/blowup_line_3.png}
@@
     \begin{definition}[Blowup of $\A^3$ at a line $L$] 
+        Given a point $p$ and a line $L$, we are interested in the unique plane
+        $H$ that contains both $p$ and $L$.  Since it contains $L$, the plane
+        can be parametrized by a point $\lambda \in \P^1$.
+
         We let the blowup-surface $\Bl_L\A^3$ be the set of all pairs
-        consisting of a point $p \in \A^3$ and a point $\lambda \in \P^1$ such
-        that $H_\lambda$ is the unique plane that contains both the point $p$
-        and the line $L$ and write
-        \begin{equation}
-            \notag
-            \Bl_L\A^3 = \left\{ (p, \lambda) \in \A^3 \times \P^1 \mid p, L \in H_\lambda \right\}.
+        consisting of a point $p \in \A^3$ and a point $\lambda \in \P^1$ that
+        parametrizes the unique plane $H$ that contains both $p$ and $L$.  We
+        write
+        \begin{equation}
+            \notag
+            \Bl_L\A^3 = \left\{ (p, \lambda) \in \A^3 \times \P^1 \mid p, L \in H \right\}.
         \end{equation}
     \end{definition}
+
+    Again, this is a fairly technical definition, so where do we go from here?
+    Assuming $L$ is parametrized by $x = at, y = bt, z = ct$, with $a, b$ and
+    $c$ non-zero, we can form the symmetric equations 
+    \begin{equation}
+        \notag
+        t = \frac{x}{a} = \frac{y}{b} = \frac{z}{c}.
+    \end{equation}
+    The line $L$ is then completely determined by two of the three possible
+    relations $xb - ya = 0$, $xc - za = 0$ and $yc - zb = 0$.  In order to find
+    the respective plane, we are interested in knowing when the plane $rx + sy
+    + tz = 0$ can be expressed in terms of the line given by $xb - ya = 0$ and
+    $yc - zb = 0$. This can be solved as a matrix problem. Given the matrix
+    \begin{equation}
+        \notag
+        A = \begin{bmatrix}
+            b & 0 & r \\
+            -a & c & s \\
+            0 & -b & t
+        \end{bmatrix}
+    \end{equation}
+    one of the columns of $A$ is linearly dependent of the two others if
+    $\det(A) = 0$. Solving $\det(A) = 0$ gives the equation
+    \begin{equation}
+        \notag
+        \det(A) = b(ct + sb) + rab = 0.
+    \end{equation}
+    We can now solve for either one of $r, s$ and $t$. Solving for $r$ we get
+    $r = -( ct + sb ) / a$ which we can substitute back into the equation for
+    our plane.  The plane is then determined by 
+    \begin{equation}
+        \notag
+        \frac{-ct - sb}{a}x + sy + tz = 0.
+    \end{equation}
+    Multiplying through by $a$ we achieve
+    \begin{align*}
+                                t(za - cx) + s(ya - bx) = 0,
+    \end{align*}
+    which gives us the restrictions that we need. The $\lambda$ that
+    parametrizes the plane $H$ is then the point $[t:s] \in \P^1$.
 
@@
         is described by the equations $x = z = 0$. Now, our goal here is to
-        find under what circumstances a plane $H_\lambda$ contains the point $p
-        = (x, y, z)$ as well as the line $L$. In other words, we want both the
+        find under what circumstances a plane $H$ contains the point $p = (x,
+        y, z)$ as well as the line $L$. In other words, we want both the
         $y$-axis and $p$ to lie in the plane $rx + sy + tz = 0$.
@@
         A plane containing the $y$-axis is then given by the equation $-tx + 0y
-        + rz = rz - tx = 0$. We can now consider the two affine charts by
-        setting $t = 1$ and $s = 1$ respectively.
+        + rz = rz - tx = 0$. The point $\lambda$ is in this case the point $[r:
+        t] \in \P^1$. We can now consider the two affine charts by setting $r =
+        1$ and $t = 1$ respectively.
 
@@
             \end{subfigure}
-            \caption{The two affine charts of the variety $\V(x^2y-z^2)$. The top
-                row is rendered by \textsc{Surfer}, and the bottom row rendered by
-                \textsc{Mathematica.} The exceptional divisor in the top row has
-                been divided out and is therefore not shown because \textsc{Surfer}
-                does a poor job at rendering planes. As seen, the charts are smooth and
-                exhibits no singularities. The intersection with the exceptional divisor is
-            rendered in blue in the bottom row.}
+            \caption{The two affine charts of the variety $\V(x^2y-z^2)$. The
+                top row is rendered by \textsc{Surfer}, and the bottom row
+                rendered by \textsc{Mathematica.} The exceptional divisor in
+                the top row has been divided out and is therefore not shown
+                because \textsc{Surfer} does a poor job at rendering planes. As
+                seen, the charts are smooth and exhibits no singularities. The
+                intersection with the exceptional divisor is rendered in blue
+                in the bottom row. The bottom row is visualized using
+            \textsc{Mathematica}.}
             \label{fig:affine_line}
@@
 
-    While in the above example we blew up a very specific line of
-    singularities, namely the $y$-axis, it can be of interest to talk a little
-    bit about how to blow up general lines. It all boils down to generalizing
-    the construction of the plane for a more general line $L$. Assuming $L$ is
-    parametrized by $x = at, y = bt, z = ct$, with $a, b$ and $c$ non-zero, we
-    can form the symmetric equations 
-    \begin{equation}
-        \notag
-        t = \frac{x}{a} = \frac{y}{b} = \frac{z}{c}.
-    \end{equation}
-    The line $L$ is then completely determined by two of the three possible
-    relations $xb - ya = 0$, $xc - za = 0$ and $yc - zb = 0$.  In order to find
-    the respective plane, we are interested in knowing when the plane $rx + sy
-    + tz = 0$ can be expressed in terms of the line given by $xb - ya = 0$ and
-    $yc - zb = 0$. This can be solved as a matrix problem. Given the matrix
-    \begin{equation}
-        \notag
-        A = \begin{bmatrix}
-            b & 0 & r \\
-            -a & c & s \\
-            0 & -b & t
-        \end{bmatrix}
-    \end{equation}
-    one of the columns of $A$ is linearly dependent of the two others if
-    $\det(A) = 0$. Solving $\det(A) = 0$ gives the equation
-    \begin{equation}
-        \notag
-        \det(A) = b(ct + sb) + rab = 0.
-    \end{equation}
-    We can now solve for either one of $r, s$ and $t$. Solving for $r$ we get
-    $r = -( ct + sb ) / a$ which we can substitute back into the equation for
-    our plane.  The plane is then determined by 
-    \begin{equation}
-        \notag
-        \frac{-ct - sb}{a}x + sy + tz = 0.
-    \end{equation}
-    Multiplying through by $a$ we achieve
-    \begin{align*}
-        (-sb - ct)x + sya + tza &= -sbx - ctx + sya + tza \\
-                                &= t(za - cx) + s(ya - bx) = 0,
-    \end{align*}
-    which gives us the restrictions that we need.
-
-    \clearpage 
-    \appendix
+    \section{Constructing singular surfaces}
+    \label{sec:constructing_singular_surfaces}
+
+    In this section we take a brief look at how one can construct singular
+    surfaces with a certain symmetry. The way we will be doing this is based on
+    the set theoretic results given in \cref{crl:set_one} and
+    \cref{crl:set_two}. This is purely exploratory and is merely a matter of
+    trying things out and seeing what the results are.
+
+    As a baseline for our surface, we wish to use the zero set of the function
+    \begin{equation*}
+        f(x, y, z) = (x - y)(x + y)(x - z)(x + z)(y - z)(y + z).
+    \end{equation*}
+    We let $\mathcal{B} = \V(f(x, y, z))$ This surface is the intersection of
+    six planes, as shown in \cref{fig:construction_one}. This surface has a
+    very nice set of singularities, namely on the intersection between the
+    planes. It also has a nice symmetry that we want our final surface to
+    inherit.
+
+    We can now take $\mathcal{B}$ and ``flesh it out'' by adding another
+    polynomial to the defining function. We want to add a surface that has a
+    group of symmetry that contains the symmetry group of $\mathcal{B}$. In our
+    case, the unit sphere seems a reasonable choice. The unit sphere is given
+    by $g(x, y, z) = x^2 + y^2 + z^2 - 1$. We let $\mathcal{S} = \V(g(x, y,
+    z))$. Define $\mathcal{B} + \mathcal{S} = \V(f(x, y, z) + g(x, y, z))$.
+    This surface has now gone from singular to smooth, but, while a bit hard to
+    see, it does retain all the symmetries of $\mathcal{B}$ as shown in
+    \cref{fig:construction_two}.
+
+    In order to regain the set of singularities, we can instead add powers of
+    $\mathcal{S}$ to $\mathcal{B}$. By powers of $\mathcal{S}$ we mean
+    $\mathcal{S}^n = \V(g(x, y, z)^n)$. If $n$ is an even number, we get a
+    surface with holes as shown in \cref{fig:construction_three}. If we instead
+    were to take $n$ an odd number, we get a solid surface with no holes, as in
+    \cref{fig:construction_four}. The higher the value we chose for $n$, the
+    more protruding the nodes of the surface become.
+
+    \begin{figure}[h]
+        \centering
+        \begin{subfigure}[t]{0.2\textwidth}
+            \includegraphics[width=\textwidth]{pictures/construction_one.png}
+            \caption{$\mathcal{B}$}
+            \label{fig:construction_one}
+        \end{subfigure}
+        ~
+        \begin{subfigure}[t]{0.2\textwidth}
+            \includegraphics[width=\textwidth]{pictures/construction_1_5.png}
+            \caption{$\mathcal{B}+ \mathcal{S}$}
+            \label{fig:construction_two}
+        \end{subfigure}
+        ~
+        \begin{subfigure}[t]{0.2\textwidth}
+            \includegraphics[width=\textwidth]{pictures/construction_two.png}
+            \caption{$\mathcal{B} + \mathcal{S}^n$ for $n$ even.}
+            \label{fig:construction_three}
+        \end{subfigure}
+        ~
+        \begin{subfigure}[t]{0.2\textwidth}
+            \includegraphics[width=\textwidth]{pictures/construction_odd.png}
+            \caption{$\mathcal{B} + \mathcal{S}^n$ for $n$ odd.}
+            \label{fig:construction_four}
+        \end{subfigure}
+        \caption{Constructing a singular surface with a given symmetry and set
+            of singularities. The surface $\mathcal{B}$ is the product of six
+            planes. The surface $\mathcal{S}$ is the unit sphere. By adding
+            various powers of $\mathcal{S}$ to $\mathcal{B}$ we achieve a
+        surface that retains the symmetries and singularities of
+    $\mathcal{B}$.}
+        \label{fig:construction}
+    \end{figure}
+
+    \section{Conclusion}
+    \label{sec:conclusion}
+
+    This paper was meant as a way of introducing the reader to some of the
+    aspects of algebraic geometry, specifically dealing with algebraic
+    surfaces. The intention was to, by restricting our attention to specific
+    examples, avoid having to take into account the fairly theoretical
+    mathematics that lie behind the scenes.  We have briefly touched upon
+    different kinds of theory related to these algebraic surfaces, which might
+    have piqued an interest.  Algebraic surfaces, while being very intuitive
+    objects in some sense, trying to figure out what they really are can
+    quickly prove a challenge because the mathematics is so involved.
+
+    Algebraic surfaces can also work as a solid entry point into visualization
+    techniques and computational mathematics, and while not talked about here,
+    we have seen some end results of the mathematical endeavours related to this
+    area of expertise, through the usage of \textsc{Surfer},
+    \textsc{Mathematica} and \textsc{Python}.
+
+    As the author of this text, it has been great fun trying to wrap my head
+    around these things, and I hope I will get the opportunity to explore this
+    field of mathematics in greater detail.  
+    \clearpage \appendix
+
+    \section{Installing visualization software}
+
+    The algebraic surface visualization industry is definitely not a million
+    dollar business, hence there are only a select few ``out of the box''
+    visualization softwares out there. Most notably are \textsc{Surf},
+    \textsc{Surf-Ex}, \textsc{Spicy} and \textsc{Surfer}. While fairly modest
+    in its features, \textsc{Surfer} is the one we have chosen for this paper,
+    due to its ease of installation and ease of use. In that regard, it is
+    perfect for this kind of introductory/exploratory text.  The software can
+    be downloaded from
+    \begin{center}
+        \url{https://imaginary.org/program/surfer#all-downloads}
+    \end{center}
+    and is compiled for use with Windows, \textsc{OS-X} and Linux operating
+    systems. Debian based Linux distributions will want to download the
+    \textsf{.deb} package, where as Red-Hat systems will want to use the
+    \textsf{.rpm} package.
+
+    If one want more options, then \textsc{Surf-Ex} is the software of choice,
+    however this has proven to be very difficult to install as it uses
+    \textsc{Surf} as a back end. In any case, the software can be found on
+    \begin{center}
+        \url{http://surfex.algebraicsurface.net} 
+    \end{center}
+    along with installation guidelines.
 
@@
     \lstset{deletekeywords={xrange}}
-    \begin{TMcode}{python}{curves.py - \url{http://stackoverflow.com/a/2484594}}{Plotting algebraic curves}
+    \begin{TMcode}{python}{curves.py -
+        \url{http://stackoverflow.com/a/2484594}}{Plotting algebraic curves}
     import matplotlib.pyplot

### 2016-08-04 project/article/src/article.tex

@@
 \newcommand{\C}{\ensuremath{\mathbb{C}}}
+\newcommand{\V}{\ensuremath{\mathbb{V}}}
 \newcommand{\proj}{\ensuremath{\mathbb{P}}}
+\renewcommand{\A}{\ensuremath{\mathbb{A}}}
 \renewcommand{\hom}{\ensuremath{h_{\mathrm{hom}}}}
@@
 
-    \begin{figure}[h]
+    \begin{figure}[]
         \centering
@@
 
-    The surfaces considered so far are over the real numbers, i.e, the
+    The surfaces considered so far are over the real numbers, i.e., the
     coefficients $a_{ijk}$ are real numbers. We can instead work over the
@@
 
-    Quite informally, projective space formalizes the notion of parallell lines
+    Quite informally, projective space formalizes the notion of parallel lines
     intersecting at infinity. In order to get a better understanding of the
@@
     of intersection in the $(x, y)$-plane. Similarly, we can look at the set of
-    points $\left[ x : y : 0 \right]$. This correponds, in $S^2$, to the circle
+    points $\left[ x : y : 0 \right]$. This corresponds, in $S^2$, to the circle
     of intersection in the $(x, y)$-plane.
@@
 
-    The general definition of projective $n$-space is completely anologous to
+    The general definition of projective $n$-space is completely analogous to
     the definition of the projective plane, the set of lines in $\C^{n+1}$
@@
     to the projective geometry introduced above.\footnote{This following
-    property can be seen directly by visualizing the homogenous polynomial
+    property can be seen directly by visualizing the homogeneous polynomial
     formed by a polynomial $g(x, y)$ and noticing that the figure does not
@@
     \end{proof}
-    Since we are working with homogenous coordinates, the condition $\hom(w, x,
+    Since we are working with homogeneous coordinates, the condition $\hom(w, x,
     y, z) = 0$ is well defined on the equivalence class discussed above,
@@
     Given a homogeneous polynomial $\hom(w, x, y, z)$ we can dehomogenize the
-    polynomial by \emph{fixing} one of the variables.
+    polynomial by \emph{fixing} one of the variables. By setting $w = f(x, y,
+    z)$ in the polynomial $\hom(w, x, y, z)$ we get a new polynomial $h(x,
+    y, z)$ which is not necessarily homogeneous.
 
@@
     Now, what do these surfaces actually look like? We give some examples that
-    is going to illustrate the above definition.
+    hopefully is going to illustrate the above definition.
 
@@
         that later.
-        \begin{figure}[h!]
+        \begin{figure}[]
             \centering
@@
         This surface has degree $d = 2$, and since all the terms have the same
-        degree the defining polynomial is homogenous, and hence the surface is
+        degree the defining polynomial is homogeneous, and hence the surface is
         invariant under scaling.  
@@
         $(1-z)z^2$ term). 
-        \begin{figure}[h!]
+        \begin{figure}[]
             \centering
-            \includegraphics[width=0.3\linewidth]{../pictures/ding_dong.png}
+            \includegraphics[width=0.2\linewidth]{../pictures/ding_dong.png}
             \caption{The ding-dong cubic given by the equation $h(x, y, z) =
@@
 
-    Another very interesting  
+    Another very interesting third-degree surface, or \emph{cubic}, is the
+    Cayley cubic. 
+    \begin{exmpl}
+        This is given by the homogeneous equation
+        \begin{equation}
+            \notag
+            \hom(w, x, y, z) = \frac{1}{w} + \frac{1}{x} + \frac{1}{y} + \frac{1}{z} = 0.
+        \end{equation}
+        In order to visualize this, we dehomogenize the polynomial by
+        multiplying through by the common denominator and setting $w = f(x, y,
+        z)$ for some polynomial $f$. It happens to be so that setting $w = 1 - x
+        - y - z$ yields a very visually pleasing surface. This surface is named
+        the \emph{Cayley cubic} after the mathematician Arthur Cayley.  Setting
+        $w$ to various functions $f$ yield different visualizations of these
+        surfaces.
+
+        The equation $xyz + (1 - x - y - z)(xy + xz + yz) = 0$ yields the
+        following surface:
+        \begin{figure}[]
+            \centering
+            \includegraphics[width=0.3\linewidth]{../pictures/cayley_cubic.png}
+            \caption{The Cayley-cubic given by the homogeneous equation $h(x,
+                y, z) = 1/w + 1 /x + 1 / y + 1/z = 0$.  This specific ''version''
+                is obtained by setting $w = 1 - x - y - z$.}
+            \label{fig:}
+        \end{figure}
+        Again, we see singularities of the surface, however this surface
+        exhibits more than just one. 
+    \end{exmpl}
+    One of the more extreme surfaces in terms of its singularities is the Barth
+    decic.
+    \begin{exmpl}[Barth decic]
+        The Barth Decic is given by the polynomial
+        \begin{multline}
+            h(x, y, z) = \\
+            8(x^2 - \varphi^4y^2)(y^2 - \varphi^4z^2)(z^2 - \varphi^4 x^2)(x^4 + y^4 +
+            z^4 - 2x^2y^2 - 2x^2z^2 - 2y^2z^2) \\ + (3 + 5\varphi)(x^2 + y^2 + z^2
+            - w^2)^2 \left[ x^2 + y^2 + z^2 - (2 - \varphi)w^2 \right]^2 w^2 \\
+            = 0, 
+        \end{multline}
+        where $\varphi$ denotes the golden ratio. Again, we dehomogenize this
+        polynomial, by setting $w = 1$. This yields the following picture.
+        \begin{figure}[]
+            \centering
+            \includegraphics[width=1.0\linewidth]{../pictures/barths_decic.png}
+            \caption{The Barth decic in all its glory. For $w = 1$ it exhibits
+            300 singularities.}
+            \label{fig:}
+        \end{figure}
+    \end{exmpl}
+
+    We have now seen examples of surfaces exhibiting so called
+    \emph{singularities}, but what are these? We discuss this in the following
+    section.
+
+    \subsection{The singularities of an algebraic surface.}
+    \label{sub:the_singularities_of_an_algebraic_surface_}
+
+    \begin{defn}
+        A \emph{singularity} of a surface is a point $(x, y, z)$ where the defining
+        polynomial $h(x, y, z)$ and all the partial derivatives of $h$ vanish
+        identically. Mathematically, $(x, y, z)$ is a singularity if $h(x, y, z) =
+        0$ and
+        \begin{align*}
+            \notag
+            \frac{\partial}{\partial x} h(x, y, z) = \frac{\partial}{\partial
+            y} h(x, y, z) = \frac{\partial}{\partial z} h(x, y, z) = 0.
+        \end{align*}
+    \end{defn}
+    One interesting question to consider is whether there is a limit to how
+    many singularities a surface of degree $d$ can have. Given a degree $d$, we
+    let $\mu(d)$ denote the maximum number of singularities possible. It
+    happens to be that $\mu(d)$ is unknown for large values of $d$.
+
+    \begin{defn}
+        Let $V = \V(f_1, ..., f_m) \subset \A^n$. A singularity, or singular point, of $V$ is a point $p$ such that
+        the Jacobian matrix
+        \begin{equation}
+            \notag
+            \begin{bmatrix}
+                \frac{\partial f_1}{\partial x_1} && \ldots && \frac{\partial f_1}{\partial x_n} \\
+                \vdots && \ddots && \vdots \\
+                \frac{\partial f_m}{\partial x_1} && \ldots && \frac{\partial f_m}{\partial x_n} \\
+            \end{bmatrix} \left( p \right)
+        \end{equation}
+        has rank strictly less than $\min(m, n)$.
+    \end{defn}
     \clearpage
